---
title: "Text Analysis Work - Activities"
author: "Noelle & Gael"
date: '2024-04-15'
output: html_document
---

# Prepare Data for Working

```{r install-packages}

# install.packages(c("tidyverse", 
#                    "tidytext",
#                    "stopwords",
#                    "widyr",
#                    "tidygraph",
#                    "tidylo",
#                    "ggraph",
#                    "wordcloud",
#                    "dplyr"))
```


```{r read-in-packages}
library(dplyr)
library(tidyverse)
library("magrittr") 
library(tidyverse)
library(tidytext)
```


```{r read-in-data}
library(readr)
journals <- read_csv("..data/journals.csv") 
```



# Figure out where Freeland goes locally most often
```{r separate-for-multiple-places}
#sometimes multiple things are in a list in one entry of a column, we want to separate to be able to count them all individually

loc_place <- journals %>%
  separate(local_place, c("place_1", "place_2", "place_3", "place_4"), sep = ",")
```
```{r reorganize-to-combine-separated-places}
loc_1 <- loc_place %>%
  select(place_1, journal, journal_entry, date_mdy, month, transcription_accuracy, location, location_accuracy, local_place_accuracy, latitude,latitude_origin, longitude, longitude_origin, quantity, unit, item, letter, letter_to, letter_from, wind_direction_am, wind_direction_pm, wind_direction_night, weather_condition_am, weather_condition_pm, weather_condition_night, temperature_am, temperature_pm, temperature_night, image_path, image_description, recorder, notes, noteworthy, second_check) %>%
  rename(place = place_1)
loc_2 <- loc_place %>%
  select(place_2, journal, journal_entry, date_mdy, month, transcription_accuracy, location, location_accuracy, local_place_accuracy, latitude,latitude_origin, longitude, longitude_origin, quantity, unit, item, letter, letter_to, letter_from, wind_direction_am, wind_direction_pm, wind_direction_night, weather_condition_am, weather_condition_pm, weather_condition_night, temperature_am, temperature_pm, temperature_night, image_path, image_description, recorder, notes, noteworthy, second_check) %>%
  rename(place = place_2)
loc_3 <- loc_place %>%
  select(place_3, journal, journal_entry, date_mdy, month, transcription_accuracy, location, location_accuracy, local_place_accuracy, latitude,latitude_origin, longitude, longitude_origin, quantity, unit, item, letter, letter_to, letter_from, wind_direction_am, wind_direction_pm, wind_direction_night, weather_condition_am, weather_condition_pm, weather_condition_night, temperature_am, temperature_pm, temperature_night, image_path, image_description, recorder, notes, noteworthy, second_check) %>%
  rename(place = place_3)
loc_4 <- loc_place %>%
  select(place_4, journal, journal_entry, date_mdy, month, transcription_accuracy, location, location_accuracy, local_place_accuracy, latitude,latitude_origin, longitude, longitude_origin, quantity, unit, item, letter, letter_to, letter_from, wind_direction_am, wind_direction_pm, wind_direction_night, weather_condition_am, weather_condition_pm, weather_condition_night, temperature_am, temperature_pm, temperature_night, image_path, image_description, recorder, notes, noteworthy, second_check) %>%
  rename(place = place_4)

loc_place <- dplyr::bind_rows(loc_1, loc_2, loc_3, loc_4)
```

```{r standardize-formatting}
loc_place <- loc_place %>%
  mutate(place = str_to_title(place)) %>%
  mutate(place = trimws(place, "both")) #str_squish 

```

```{r count-places}
loc_count <- loc_place %>%
    count(place, sort = TRUE)
```
```{r print-place-count}
print(loc_count)
```

```{r create-bar-chart-for-places}
loc_count %>%
    filter(n > 30) %>%
    filter(n < 2000) %>% #2000 to get rid of NA (may change)
  ggplot(aes(x=place, y=n)) + 
   geom_bar(stat = "identity")+
    theme(axis.text.x = element_text(angle = 60, hjust = 1))+
   labs(y = "n", x = "Location", title = "Local Locations in Freeland Bunker's Journals")
```


# Frequent words in association with Home
```{r filter-for-home}
library(stringr)

Home <-  dplyr::filter(journals, grepl('Home| home| Home', local_place))

#https://stackoverflow.com/questions/22850026/filter-rows-which-contain-a-certain-string
```
```{r separate-out-words-in-entries}
(journal_words <- Home %>%
    unnest_tokens(word, journal_entry) %>%
    count(journal, word, sort = TRUE) %>%
    filter(!word %in% stop_words$word))
```
```{r find-tfidf}
journal_tf_idf <- journal_words %>%
    bind_tf_idf(word, journal, n) %>%
    arrange(desc(tf_idf))
```
```{r plot-tfidf}
journal_tf_idf %>%
#  filter(tf_idf >= 0.02) %>%
    group_by(journal) %>%
    slice_max(tf_idf, n = 10) %>%
    ggplot(aes(tf_idf, fct_reorder(word, tf_idf), fill = journal)) +
    geom_col(show.legend = FALSE, width = 0.5) +
    facet_wrap(vars(journal), scales = "free") +
    scale_x_continuous(expand = c(0,0)) +
    labs(y = NULL, x = "tf-idf", title = "Highest tf-idf words in Freeland Bunker's Journal entries associated with mention of being at home")
```

# Frequent words in association with the Lodge of Good Templars
```{r filter-for-lodge}
library(stringr)

LoGT <-  dplyr::filter(journals, grepl('Lodge of Good Templars| Lodge of Good Templars| Lodge of  Good Templars|	Lodge of the Good Templars', local_place))

#https://stackoverflow.com/questions/22850026/filter-rows-which-contain-a-certain-string
```
```{r separate-out-words-in-entries}
(journal_words <- LoGT %>%
    unnest_tokens(word, journal_entry) %>%
    count(journal, word, sort = TRUE) %>%
    filter(!word %in% stop_words$word))
```
```{r find-tfidf}
journal_tf_idf <- journal_words %>%
    bind_tf_idf(word, journal, n) %>%
    arrange(desc(tf_idf))
```
```{r plot-tfidf}
journal_tf_idf %>%
  filter(tf_idf >= 0.02) %>%
    group_by(journal) %>%
    slice_max(tf_idf, n = 10) %>%
    ggplot(aes(tf_idf, fct_reorder(word, tf_idf), fill = journal)) +
    geom_col(show.legend = FALSE) +
    facet_wrap(vars(journal), scales = "free") +
    scale_x_continuous(expand = c(0,0)) +
    labs(y = NULL, x = "tf-idf", title = "Highest tf-idf words in Freeland Bunker's Journal entries associated with mention of being at The Lodge of Good Templars")
```


# Frequent words in association with Hall
```{r filter-for-Hall}
library(stringr)

Hall <-  dplyr::filter(journals, grepl('Hall|Hall | Hall| hall', local_place))

#https://stackoverflow.com/questions/22850026/filter-rows-which-contain-a-certain-string
```
```{r separate-out-words-in-entries}
(journal_words <- Hall %>%
    unnest_tokens(word, journal_entry) %>%
    count(journal, word, sort = TRUE) %>%
    filter(!word %in% stop_words$word))
```
```{r find-tfidf}
journal_tf_idf <- journal_words %>%
    bind_tf_idf(word, journal, n) %>%
    arrange(desc(tf_idf))
```
```{r plot-tfidf}
journal_tf_idf %>%
  filter(tf_idf >= 0.02) %>%
    group_by(journal) %>%
    slice_max(tf_idf, n = 10) %>%
    ggplot(aes(tf_idf, fct_reorder(word, tf_idf), fill = journal)) +
    geom_col(show.legend = FALSE) +
    facet_wrap(vars(journal), scales = "free") +
    scale_x_continuous(expand = c(0,0)) +
    labs(y = NULL, x = "tf-idf", title = "Highest tf-idf words in Freeland Bunker's Journal entries associated with mention of being at The Hall")
```

# Frequent words in association with Shop
```{r filter-for-Shop}
Shop <-  dplyr::filter(journals, grepl('Shop|Shop | Shop| shop', local_place))

#https://stackoverflow.com/questions/22850026/filter-rows-which-contain-a-certain-string
```
```{r separate-out-words-in-entries}
(journal_words <- Shop %>%
    unnest_tokens(word, journal_entry) %>%
    count(journal, word, sort = TRUE) %>%
    filter(!word %in% stop_words$word))
```
```{r find-tfidf}
journal_tf_idf <- journal_words %>%
    bind_tf_idf(word, journal, n) %>%
    arrange(desc(tf_idf))
```
```{r plot-tfidf}
journal_tf_idf %>%
  filter(tf_idf >= 0.02) %>%
    group_by(journal) %>%
    slice_max(tf_idf, n = 10) %>%
    ggplot(aes(tf_idf, fct_reorder(word, tf_idf), fill = journal)) +
    geom_col(show.legend = FALSE) +
    facet_wrap(vars(journal), scales = "free") +
    scale_x_continuous(expand = c(0,0)) +
    labs(y = NULL, x = "tf-idf", title = "Highest tf-idf words in Freeland Bunker's Journal entries associated with mention of being at The Shop")
```

# Frequent words in association with Wharf
```{r filter-for-Wharf}
Wharf <-  dplyr::filter(journals, grepl('Wharf|Wharf | Wharf| wharf', local_place))

#https://stackoverflow.com/questions/22850026/filter-rows-which-contain-a-certain-string
```
```{r separate-out-words-in-entries}
(journal_words <- Wharf %>%
    unnest_tokens(word, journal_entry) %>%
    count(journal, word, sort = TRUE) %>%
    filter(!word %in% stop_words$word))
```
```{r find-tfidf}
journal_tf_idf <- journal_words %>%
    bind_tf_idf(word, journal, n) %>%
    arrange(desc(tf_idf))
```
```{r plot-tfidf}
journal_tf_idf %>%
  filter(tf_idf >= 0.02) %>%
    group_by(journal) %>%
    slice_max(tf_idf, n = 10) %>%
    ggplot(aes(tf_idf, fct_reorder(word, tf_idf), fill = journal)) +
    geom_col(show.legend = FALSE) +
    facet_wrap(vars(journal), scales = "free") +
    scale_x_continuous(expand = c(0,0)) +
    labs(y = NULL, x = "tf-idf", title = "Highest tf-idf words in Freeland Bunker's Journal entries associated with mention of being at The Wharf")
```


# Distribution of local place mentions across journals
```{r separate-multiple-places}
loc_place <- journals %>%
  select(journal, local_place, location) %>%
  separate(local_place, c("place_1", "place_2", "place_3", "place_4"), sep = ",")
```
```{r reorganize-to-combine-separated-places}
loc_1 <- loc_place %>%
  select(place_1, journal) %>%
  rename(place = place_1)
loc_2 <- loc_place %>%
  select(place_2, journal) %>%
  rename(place = place_2)
loc_3 <- loc_place %>%
  select(place_3, journal) %>%
  rename(place = place_3)
loc_4 <- loc_place %>%
  select(place_4, journal) %>%
  rename(place = place_4)

loc_count <- dplyr::bind_rows(loc_1, loc_2, loc_3, loc_4)
```

```{r standardize-formatting}
loc_count <- loc_count %>%
  mutate(place = str_to_title(place)) %>%
  mutate(place = trimws(place, "both"))  

```

```{r count-and-filter-each-journal}
loc_count_1 <- loc_count %>%
    filter(journal == 1) %>%
           count(place) 
loc_count_2 <- loc_count %>%
    filter(journal == 2) %>%
           count(place) 
loc_count_3 <- loc_count %>%
    filter(journal == 3) %>%
           count(place) 
loc_count_4 <- loc_count %>%
    filter(journal == 4) %>%
           count(place) 
loc_count_5 <- loc_count %>%
    filter(journal == 5) %>%
           count(place) 
loc_count_6 <- loc_count %>%
    filter(journal == 6) %>%
           count(place) 
loc_count_9 <- loc_count %>%
    filter(journal == 9) %>%
           count(place) 
loc_count_12 <- loc_count %>%
    filter(journal == 12) %>%
           count(place)
loc_count_13 <- loc_count %>%
    filter(journal == 13) %>%
           count(place)
loc_count_15 <- loc_count %>%
    filter(journal == 15) %>%
           count(place) 

```
```{r assign-new-column}
loc_count_1$journal <- 1
loc_count_2$journal <- 2
loc_count_3$journal <- 3
loc_count_4$journal <- 4
loc_count_5$journal <- 5
loc_count_6$journal <- 6
loc_count_9$journal <- 9
loc_count_12$journal <- 12
loc_count_13$journal <- 13
loc_count_15$journal <- 15
```

```{r rejoin-data}

loc_count <- dplyr::bind_rows(loc_count_1, loc_count_2, loc_count_3, loc_count_4, loc_count_5, loc_count_6, loc_count_9, loc_count_12,loc_count_13, loc_count_15)
```

```{r filter-for-top-places}
loc_count <- loc_count %>%
filter(place %in% c("Home", "Shop", "Hall", "Wharf"))

```

```{r bar-chart-for-place-distribution}
loc_count %>%
  group_by(journal) %>%
 # filter(n < 500) %>%  #include home & Shop
 #  filter(n < 40) %>%
 #  filter(n > 5) %>%
  #  filter(word != NA) %>%
  ggplot(aes(x=place, y=n)) + 
   geom_bar(stat = "identity")+
   facet_wrap(vars(journal))+
    theme(axis.text.x = element_text(angle = 60, hjust = 1))+
   labs(y = "n", x = "Location", title = "Distribution of local place mentions across journals")
```

## Prepare data for mapping
Add in latitude and longitude

#local places
```{r create-mapping-data-set}
place_map <-   loc_place %>%
  filter(place == c('Home','Shop','Lodge Of Good Templars','Lodge Of The Good Templars','Hall','Wharf' ))

# dplyr::filter(loc_place, grepl('Home|Shop|Lodge Of Good Templars|Lodge Of The Good Templars|Hall|Wharf', place))
```
```{r assign-latitudes}
mapping <- c("Home" = 44.39044, 
"Shop" = 44.39078, #Double Check
"Lodge Of Good Templars" = 0,
"Lodge Of The Good Templars" = 0,
"Hall" = 44.39168,
"Wharf" = 44.38937)

place_map$latitude <- mapping[as.character(place_map$place)]
```
```{r assign-longitude}
mapping <- c("Home" = -68.09314, 
"Shop" = -68.09316,
"Lodge Of Good Templars" = 0,
"Lodge Of The Good Templars" = 0,
"Hall" = -68.09103,
"Wharf" = -68.08918)

place_map$longitude <- mapping[as.character(place_map$place)]
```
```{r get-rid-of-NA}
place_map <- place_map %>%
  filter(latitude != "NA")
```
```{r export-dataset}
write.csv(place_map, "places.csv", row.names=FALSE)
```


#local people
Top mentions of Freeland visiting People's homes in local_place
- Abner Tracy's
- Joseph Bunker's
- James Bourgett's
- T. R. Hammond's House
- N. A. Bickford's
- Thomas H. Smallidge's (Thomas Smallidge's | Thomas Smallige's)

