---
title: "Text Analysis Work for letters"
author: "Noelle & Gael"
date: '2024-04-15'
output: html_document
---
```{r install-packages}
#install.packages(c("tidyverse", 
 #                  "tidytext",
  #                 "stopwords",
   #                "widyr",
    #               "tidygraph",
     #              "tidylo",
      #             "ggraph",
       #            "wordcloud",
        #           "dplyr"))
```


```{r read-in-packages}
library(dplyr)
library(tidyverse)
library(magrittr) 
library(tidyverse)
library(tidytext)
```



```{r read-in-data}
library(readr)
journals <- read_csv("..data/journals.csv") 
```


# Create new dataframe for letters

```{r filter-out-no-letter-entries}
letters <- journals %>%
  filter(letter != "no letter")
```

```{r filter-for-read-letters}
read_letters <- letters %>%
  filter(letter_from != "NA")
  
```
```{r filter-for-written-letters}
write_letters <- letters %>%
  filter(letter_to != "NA")
```


# Count people Freeland most often recived letters from

```{r separate-for-multiple-people}
#sometimes multiple things are in a list in one entry of a column, we want to separate to be able to count them all individually

read_letters <- read_letters %>%
  separate(letter_from, c("person_1", "person_2", "person_3", "person_4"), sep = ",")
```

```{r reorganize-to-combine-separated-people}
per_1 <- read_letters %>%
  select(person_1, journal, journal_entry, date_mdy, month, transcription_accuracy, location, location_accuracy, local_place, local_place_accuracy, latitude,latitude_origin, longitude, longitude_origin, quantity, unit, item, letter, wind_direction_am, wind_direction_pm, wind_direction_night, weather_condition_am, weather_condition_pm, weather_condition_night, temperature_am, temperature_pm, temperature_night, image_path, image_description, recorder, notes, noteworthy, second_check) %>%
  rename(from = person_1)
per_2 <- read_letters %>%
  select(person_2, journal, journal_entry, date_mdy, month, transcription_accuracy, location, location_accuracy, local_place, local_place_accuracy, latitude,latitude_origin, longitude, longitude_origin, quantity, unit, item, letter, wind_direction_am, wind_direction_pm, wind_direction_night, weather_condition_am, weather_condition_pm, weather_condition_night, temperature_am, temperature_pm, temperature_night, image_path, image_description, recorder, notes, noteworthy, second_check) %>%
  rename(from = person_2)
per_3 <- read_letters %>%
  select(person_3, journal, journal_entry, date_mdy, month, transcription_accuracy, location, location_accuracy, local_place, local_place_accuracy, latitude,latitude_origin, longitude, longitude_origin, quantity, unit, item, letter, wind_direction_am, wind_direction_pm, wind_direction_night, weather_condition_am, weather_condition_pm, weather_condition_night, temperature_am, temperature_pm, temperature_night, image_path, image_description, recorder, notes, noteworthy, second_check) %>%
  rename(from = person_3)
per_4 <- read_letters %>%
  select(person_4, journal, journal_entry, date_mdy, month, transcription_accuracy, location, location_accuracy, local_place, local_place_accuracy, latitude,latitude_origin, longitude, longitude_origin, quantity, unit, item, letter, wind_direction_am, wind_direction_pm, wind_direction_night, weather_condition_am, weather_condition_pm, weather_condition_night, temperature_am, temperature_pm, temperature_night, image_path, image_description, recorder, notes, noteworthy, second_check) %>%
  rename(from = person_4)

read_letters <- dplyr::bind_rows(per_1, per_2, per_3, per_4)
```
```{r standardize-formatting}
read_letters <- read_letters %>%
  mutate(from = str_to_title(from)) %>%
  mutate(from = trimws(from, "both")) %>%
  filter(from != "NA")
```

```{r count-read-letters}
(read_count <- read_letters %>%
   filter(from!= "NA") %>%
    count(from, sort = TRUE))
```


# Count people Freeland most often wrote letters to

```{r separate-for-multiple-people}
write_letters <- write_letters %>%
  separate(letter_to, c("person_1", "person_2", "person_3", "person_4"), sep = ",")
```

```{r reorganize-to-combine-separated-people}
per_1 <- write_letters %>%
  select(person_1, journal, journal_entry, date_mdy, month, transcription_accuracy, location, location_accuracy, local_place, local_place_accuracy, latitude,latitude_origin, longitude, longitude_origin, quantity, unit, item, letter, wind_direction_am, wind_direction_pm, wind_direction_night, weather_condition_am, weather_condition_pm, weather_condition_night, temperature_am, temperature_pm, temperature_night, image_path, image_description, recorder, notes, noteworthy, second_check) %>%
  rename(to = person_1)
per_2 <- write_letters %>%
  select(person_2, journal, journal_entry, date_mdy, month, transcription_accuracy, location, location_accuracy, local_place, local_place_accuracy, latitude,latitude_origin, longitude, longitude_origin, quantity, unit, item, letter, wind_direction_am, wind_direction_pm, wind_direction_night, weather_condition_am, weather_condition_pm, weather_condition_night, temperature_am, temperature_pm, temperature_night, image_path, image_description, recorder, notes, noteworthy, second_check) %>%
  rename(to = person_2)
per_3 <- write_letters %>%
  select(person_3, journal, journal_entry, date_mdy, month, transcription_accuracy, location, location_accuracy, local_place, local_place_accuracy, latitude,latitude_origin, longitude, longitude_origin, quantity, unit, item, letter, wind_direction_am, wind_direction_pm, wind_direction_night, weather_condition_am, weather_condition_pm, weather_condition_night, temperature_am, temperature_pm, temperature_night, image_path, image_description, recorder, notes, noteworthy, second_check) %>%
  rename(to = person_3)
per_4 <- write_letters %>%
  select(person_4, journal, journal_entry, date_mdy, month, transcription_accuracy, location, location_accuracy, local_place, local_place_accuracy, latitude,latitude_origin, longitude, longitude_origin, quantity, unit, item, letter, wind_direction_am, wind_direction_pm, wind_direction_night, weather_condition_am, weather_condition_pm, weather_condition_night, temperature_am, temperature_pm, temperature_night, image_path, image_description, recorder, notes, noteworthy, second_check) %>%
  rename(to = person_4)

write_letters <- dplyr::bind_rows(per_1, per_2, per_3, per_4)
```
```{r standardize-formatting}
write_letters <- write_letters %>%
  mutate(to = str_to_title(to)) %>%
  mutate(to = trimws(to, "both")) %>%
  filter(to != "NA")
```


```{r count-written-letters}
(write_count <- write_letters %>%
   filter(to != "NA") %>%
    count(to, sort = TRUE))

```

#Info on idividuals
We picked top five letter writers/recievers & read any journal entry that mentioned them.

```{r filter-for-Kittridge}
ben_kitt <-letters %>%
  filter(grepl('Benjamin Kittridge|benjamin kittridge', letter_from))
ben_kitt_2 <-letters %>%
  filter(grepl('Benjamin Kittridge|benjamin kittridge', letter_to))

ben_kitt <- dplyr::bind_rows(ben_kitt, ben_kitt_2)
```
```{r find-words-mentioned-most-with-Kittridge}
(BK_words <- ben_kitt %>%
    unnest_tokens(word, journal_entry) %>%
    count(journal, word, sort = TRUE) %>%
    filter(!word %in% stop_words$word))
```
Bk mentions outside of letters
```{r filter-for-Kittridge}
bk_mention <-journals %>%
  filter(grepl('Benj. Kittridge|Benjamin Kittredge|Kittredge', journal_entry)) %>%
  filter(letter == "no letter")
```
```{r find-words-mentioned-most-with-Kittridge}
(bk_word_xletter <- bk_mention %>%
  filter(letter == "no letter") %>%
    unnest_tokens(word, journal_entry) %>%
    count(journal, word, sort = TRUE) %>%
    filter(!word %in% stop_words$word))
```
Info Obtained:
- Benjamin Kittridge lived in Bar Harbor
- talks to Freeland about a vessel
- gets meal bushels from Freeland
- Only visited in 1873

#CC Burrill
```{r filter-for-Burrill}
CC_Burr <-letters %>%
  filter(grepl('Charles C. Burrill|C. C. Burrill', letter_from))
CC_Burr_2 <-letters %>%
  filter(grepl('Charles C. Burrill|C. C. Burrill', letter_to))

CC_Burr <- dplyr::bind_rows(CC_Burr, CC_Burr_2)
```
```{r find-words-mentioned-most-with-Burrill}
(CC_words <- CC_Burr %>%
    unnest_tokens(word, journal_entry) %>%
    count(journal, word, sort = TRUE) %>%
    filter(!word %in% stop_words$word))
```
CB mentions outside of letters
```{r filter-for-Burrill}
CB_mention <-journals %>%
  filter(grepl('CC Burrill|C.C. Burrill|C. C. Burrill|C C Burrill| Charles C. Burrill| Charles C Burrill|Chas. C. Burrill', journal_entry)) %>%
  filter(letter == "no letter")
```
```{r find-words-mentioned-most-with-Burrill}
(cb_word_xletter <- CB_mention %>%
  filter(letter == "no letter") %>%
    unnest_tokens(word, journal_entry) %>%
    count(journal, word, sort = TRUE) %>%
    filter(!word %in% stop_words$word))
```
Info Obtained:
- Work with Freeland on life insurance
- Had dinner with Freeland on occation
- Travelled together to pospect Harbor
- Lived in Ellsworth
- Only mentioned visiting in 1871-1873

#Thomas Bunker
```{r filter-for-Bunker}
Tom_Bunk <-letters %>%
  filter(grepl('Thomas Bunker', letter_from))
Tom_Bunk_2 <-letters %>%
  filter(grepl('Thomas Bunker', letter_to))

Tom_Bunk <- dplyr::bind_rows(Tom_Bunk, Tom_Bunk_2)
```
```{r find-words-mentioned-most-with-Bunker}
(TB_words <- Tom_Bunk %>%
    unnest_tokens(word, journal_entry) %>%
    count(journal, word, sort = TRUE) %>%
    filter(!word %in% stop_words$word))
```
TB mentions outside of letters
```{r filter-for-Bunker}
tb_mention <-journals %>%
  filter(grepl('Thomas Bunker|Thom Bunker|thomas Bunker', journal_entry)) %>%
  filter(letter == "no letter")
```
```{r find-words-mentioned-most-with-Bunker}
(tb_word_xletter <- tb_mention %>%
  filter(letter == "no letter") %>%
    unnest_tokens(word, journal_entry) %>%
    count(journal, word, sort = TRUE) %>%
    filter(!word %in% stop_words$word))
```
Info Obtained:
- Lived on the Cranberry Isles
- Freeland often visited, had meals, & stayed overnight with TB
- Used Freeland's schooner the Sea Flower, bought from FB in 1875
- mentioned visits from 1873-1877

#J.F. Kingsley
not a top letter person, but mentioned elsewhere
```{r filter-for-kingsley}
JF_King <-letters %>%
  filter(grepl('J.F. Kingsley|J.F.K|JFK|Jfk|J F Kingsley| J. F. Kingsley', letter_from))
JF_King_2 <-letters %>%
  filter(grepl('J.F. Kingsley|J.F.K|JFK|Jfk|J F Kingsley| J. F. Kingsley', letter_to))

JF_King <- dplyr::bind_rows(JF_King, JF_King_2)
```
```{r find-words-mentioned-most-with-Kingsley}
(JFK_words <- JF_King %>%
    unnest_tokens(word, journal_entry) %>%
    count(journal, word, sort = TRUE) %>%
    filter(!word %in% stop_words$word))
```
JFK mentions outside of letters
```{r filter-for-kingsley}
jfk_mention <-journals %>%
  filter(grepl('J.F. Kingsley|J.F.K|JFK|Jfk|J F Kingsley| J. F. Kingsley', journal_entry)) 
```
```{r find-words-mentioned-most-with-Kingsley}
(jfk_word_xletter <- jfk_mention %>%
    unnest_tokens(word, journal_entry) %>%
    count(journal, word, sort = TRUE) %>%
    filter(!word %in% stop_words$word))
```


# Find distribution of people read/wrote letters across journals

#Wrote

```{r filter-journals-and-count}
count_1 <- write_letters %>%
    filter(journal == 1) %>%
           count(to) 
count_2 <- write_letters %>%
    filter(journal == 2) %>%
           count(to) 
count_3 <- write_letters %>%
    filter(journal == 3) %>%
           count(to)
count_4 <- write_letters %>%
    filter(journal == 4) %>%
           count(to)
count_5 <- write_letters %>%
    filter(journal == 5) %>%
           count(to)
count_6 <- write_letters %>%
    filter(journal == 6) %>%
           count(to) 
count_9 <- write_letters %>%
    filter(journal == 9) %>%
           count(to)
count_12 <- write_letters %>%
    filter(journal == 12) %>%
           count(to)
count_13 <- write_letters %>%
    filter(journal == 13) %>%
           count(to)
count_15 <- write_letters %>%
    filter(journal == 15) %>%
           count(to) 

```
```{r assign-new-column}
count_1$journal <- 1
count_2$journal <- 2
count_3$journal <- 3
count_4$journal <- 4
count_5$journal <- 5
count_6$journal <- 6
count_9$journal <- 9
count_12$journal <- 12
count_13$journal <- 13
count_15$journal <- 15
```

```{r join-data}

to_count <- dplyr::bind_rows(count_1, count_2, count_3, count_4, count_5, count_6, count_9, count_12,count_13, count_15)
```


```{r bar-chart-distribution-of-letters}
to_count %>%
  group_by(journal) %>%
   filter(n > 3) %>%
  ggplot(aes(x=to, y=n)) + 
   geom_bar(stat = "identity")+
   facet_wrap(vars(journal))+
    theme(axis.text.x = element_text(angle = 60, hjust = 1))+
   labs(y = "n", x = "Location", title = "Distribution of letters written across journals")
```

#Read

```{r filter-journals-and-count}
count_1 <- read_letters %>%
    filter(journal == 1) %>%
           count(from) 
count_2 <- read_letters %>%
    filter(journal == 2) %>%
           count(from) 
count_3 <- read_letters %>%
    filter(journal == 3) %>%
           count(from)
count_4 <- read_letters %>%
    filter(journal == 4) %>%
           count(from)
count_5 <- read_letters %>%
    filter(journal == 5) %>%
           count(from)
count_6 <- read_letters %>%
    filter(journal == 6) %>%
           count(from) 
count_9 <- read_letters %>%
    filter(journal == 9) %>%
           count(from)
count_12 <- read_letters %>%
    filter(journal == 12) %>%
           count(from)
count_13 <- read_letters %>%
    filter(journal == 13) %>%
           count(from)
count_15 <- read_letters %>%
    filter(journal == 15) %>%
           count(from) 

```
```{r assign-new-column}
count_1$journal <- 1
count_2$journal <- 2
count_3$journal <- 3
count_4$journal <- 4
count_5$journal <- 5
count_6$journal <- 6
count_9$journal <- 9
count_12$journal <- 12
count_13$journal <- 13
count_15$journal <- 15
```

```{r join-data}

from_count <- dplyr::bind_rows(count_1, count_2, count_3, count_4, count_5, count_6, count_9, count_12,count_13, count_15)
```


```{r bar-chart-distribution-of-letters}
from_count %>%
  group_by(journal) %>%
   filter(n > 3) %>%
  ggplot(aes(x=from, y=n)) + 
   geom_bar(stat = "identity")+
   facet_wrap(vars(journal))+
    theme(axis.text.x = element_text(angle = 60, hjust = 1))+
   labs(y = "n", x = "Location", title = "Distribution of letters recieved across journals")
```

# writing & reading across journals
```{r filter-journals-and-count}
count_1 <- letters %>%
    filter(journal == 1) %>%
           count(letter) 
count_2 <- letters %>%
    filter(journal == 2) %>%
           count(letter) 
count_3 <- letters %>%
    filter(journal == 3) %>%
           count(letter)
count_4 <- letters %>%
    filter(journal == 4) %>%
           count(letter)
count_5 <- letters %>%
    filter(journal == 5) %>%
           count(letter)
count_6 <- letters %>%
    filter(journal == 6) %>%
           count(letter) 
count_9 <- letters %>%
    filter(journal == 9) %>%
           count(letter)
count_12 <- letters %>%
    filter(journal == 12) %>%
           count(letter)
count_13 <- letters %>%
    filter(journal == 13) %>%
           count(letter)
count_15 <- letters %>%
    filter(journal == 15) %>%
           count(letter) 

```
```{r assign-new-column}
count_1$journal <- 1
count_2$journal <- 2
count_3$journal <- 3
count_4$journal <- 4
count_5$journal <- 5
count_6$journal <- 6
count_9$journal <- 9
count_12$journal <- 12
count_13$journal <- 13
count_15$journal <- 15
```

```{r join-data}

letter_count <- dplyr::bind_rows(count_1, count_2, count_3, count_4, count_5, count_6, count_9, count_12,count_13, count_15)
```


```{r bar-chart-distribution-of-letters}
letter_count %>%
  group_by(journal) %>%
 #  filter(n > 3) %>%
  ggplot(aes(x=letter, y=n)) + 
   geom_bar(stat = "identity")+
   facet_wrap(vars(journal))+
    theme(axis.text.x = element_text(angle = 60, hjust = 1))+
   labs(y = "n", x = "Letter mentioned", title = "Distribution of letters written/recieved across journals")
```


#distribution of top people
```{r filter-for-top-people}
top_from <- dplyr::filter(from_count, grepl('C. C. Burill|C.C. Burrill|C.c. Burrill|CC Burrill|Charles C. Burrill|Charles C Burill|Benjamin Kittridge|Thomas Bunker|J.F. Kingsley|J.F.K|JFK|Jfk|J F Kingsley| J. F. Kingsley', from))

```

```{r bar-chart-distribution-of-letters}
top_from %>%
  group_by(journal) %>%
  ggplot(aes(x=from, y=n)) + 
   geom_bar(stat = "identity")+
   facet_wrap(vars(journal))+
    theme(axis.text.x = element_text(angle = 60, hjust = 1))+
   labs(y = "n", x = "People Receiving or Writing Letters", title = "Distribution of letters recieved across journals")
```


```{r make-name-variation-all-the-same}
mapping <- c("Capt. Thomas Bunker" = "Thomas Bunker", "Stephen Southard Thomas Bunker" = "Thomas Bunker", "Thomas Bunker" = "Thomas Bunker", "Charles C. Burrill" = "Charles C. Burrill",  "Benjamin Kittridge" = "Benjamin Kittridge")

top_to <- dplyr::filter(to_count, grepl('C. C. Burill|C.C. Burrill|C.c. Burrill|CC Burrill|Charles C. Burrill|Charles C Burill|Benjamin Kittridge|Thomas Bunker|Capt. Thomas Bunker|J.F. Kingsley|J.F.K|JFK|Jfk|J F Kingsley| J. F. Kingsley', to))
  
top_to$to <- mapping[as.character(top_to$to)]
  
top_to <- top_to %>%  
  filter(to != "NA")
  
```

```{r bar-chart-distribution-of-letters}
top_to %>%
  group_by(journal) %>%
  ggplot(aes(x=to, y=n)) + 
   geom_bar(stat = "identity")+
   facet_wrap(vars(journal))+
    theme(axis.text.x = element_text(angle = 60, hjust = 1))+
   labs(y = "n", x = "Location", title = "Distribution of letters recieved across journals")
```

##Adding in coordinates for GIS Mapping

#where reciever was
```{r create-mapping-dataset}
#https://forum.posit.co/t/creating-a-new-column-based-on-another-column/109464/6

Top_2 <-   
  dplyr::filter(write_letters, grepl('C. C. Burill|C.C. Burrill|C.c. Burrill|CC Burrill|Charles C. Burrill|Charles C Burill|Benjamin Kittridge|Thomas Bunker|Capt. Thomas Bunker|J.F. Kingsley|J.F.K|JFK|Jfk|J F Kingsley| J. F. Kingsley', to))%>%
  rename(person = to)
Top_f <-
  dplyr::filter(read_letters, grepl('C. C. Burill|C.C. Burrill|C.c. Burrill|CC Burrill|Charles C. Burrill|Charles C Burill|Benjamin Kittridge|Thomas Bunker|Capt. Thomas Bunker|J.F. Kingsley|J.F.K|JFK|Jfk|J F Kingsley| J. F. Kingsley', from))%>%
  rename(person = from)

top <- dplyr::bind_rows(Top_2, Top_f)
```
```{r assign-latitude}
mapping <- c("Charles C. Burrill" = 44.5434, "Benjamin Kittridge" = 44.387699, "Thomas Bunker" = 44.2534, "Capt Thomas Bunker" = 44.2534, "Stephen Southard Thomas Bunker" = 44.2534)

top$latitude_to <- mapping[as.character(top$person)]
```
```{r assign-longitude}
mapping <- c("Charles C. Burrill" = -68.4195, "Benjamin Kittridge" = -68.212219, "Thomas Bunker" = -68.1837, "Capt Thomas Bunker" = -68.1837, "Stephen Southard Thomas Bunker" = -68.1837)

top$longitude_to <- mapping[as.character(top$person)]
```

#where Freeland was
```{r assign-latitude}
mapping <- c("Winter Harbor" = 44.3948,                     
"West Gouldsboro, Winter Harbor" = 44.3948, 
# "NA"
"West Gouldsboro" = 44.4695,       
"Winter Harbor, Mark Island, Winter Harbor" = 44.3948,
"Hog Island, Jones Cove" = 44.4687,      
# "Salem"                                    
"Boston"  = 42.3601,                                 
"Stave Island, Winter Harbor" = 44.3948,          
"Prospect Harbor" = 44.4033,                       
"Ellsworth" = 44.5434,                  
"Bangor" = 44.8016,        
"Winter Harbor, West Gouldsboro" = 44.3948,      
"Winter Harbor, Turtle Isle, Winter Harbor" = 44.3948,
"Winter Harbor, Iron Bound, Winter Harbor" = 44.3948,
"Hancock, Petit Manan, Winter Harbor" = 44.3948, 
"Deer Isle, Bass Harbor, Winter Harbor"  = 44.3948,  
"Duck Islands, Winter Harbor" = 44.3948,
"Bass Habor, Winter Harbor"= 44.3948
# "Albany Railroad Wharf"  
)
top$latitude <- mapping[as.character(top$location)]
```
```{r assign-longitude}
mapping <- c("Winter Harbor" = -68.0830,                     
"West Gouldsboro, Winter Harbor"= -68.0830,  
# "NA"
"West Gouldsboro" = -68.0903,               
"Winter Harbor, Mark Island, Winter Harbor" = -68.0830,
"Hog Island, Jones Cove" = -68.1017,
# "Salem"                                    
"Boston" = -71.0589,                               
"Stave Island, Winter Harbor" = -68.0830,           
"Prospect Harbor" = 68.0254,                      
"Ellsworth" = -68.4195,                   
"Bangor" = -68.7712,                       
"Winter Harbor, West Gouldsboro" = -68.0830,
"Winter Harbor, Turtle Isle, Winter Harbor" = -68.0830,
"Winter Harbor, Iron Bound, Winter Harbor" = -68.0830,
"Hancock, Petit Manan, Winter Harbor" = -68.0830,
"Deer Isle, Bass Harbor, Winter Harbor" = -68.0830,
"Duck Islands, Winter Harbor" = -68.0830, 
"Bass Habor, Winter Harbor" = -68.0830  
# "Albany Railroad Wharf"  
)

top$longitude <- mapping[as.character(top$location)]
```

```{r export-data}
write.csv(top, "letter.csv", row.names=FALSE)
```

