---
title: "Wind Speed Beaufort Analysis"
output: html_document
---

## Read in Wind Data from WindData script

```{r wind-speed-patterns}
# Wind speed patterns 

wind_speed_patterns <- "(?i)\\b(strong|powerful|heavy|intense|light|gentle|soft|calm|still|moderate|steady|gusty|blustery|variable|baffling|fresh|squally|brisk|stormy|tempestuous|strong|powerful|heavy|intense|light|gentle|soft|calm|still|moderate|steady|gusty|blustery|variable|calm|pleasant|fine|clear|light|fresh breeze|scant|fine|clam|clear|light|cleared of wind|smooth|baffling wind|eased|light air|fesh|dead calm|died out|hair|no wind|becalmed|not wind enough|baffling|becalmed|cleared up|very light air of wind|light air|perfectly calm|hazed up|wind out|hauling little|baffling|fresh|squally|brisk|stormy|tempestuous|moderate|fresh breeze|strong breeze|smart breeze|fresh blow good breeze|very fresh breeze|moderating|blowy|blowing|variable|breezing|whole sail breeze|changeable|fresh trade winds|high|peak breeze|mederate|fine|perfect|hauled down|steady|fine|wind increasing|breezed up|increasing|wind out|baffling and moderate|blustering|heavy blow|heavy|strong|rough|windstorm |big storm |heavy breeze|gale|stronger|tremendous|squall|hard|blowing a gale|hauled|squalls|hauling|terrific storm |blowing hard|squalls|perfect gale|squall|high winds|blew very heavy|fearfuly|blowed fearfuly|hauling out|rain squalls|growing rough|very moderate, calm|fresh breeze, rough|heavy blow, rough|sair|mode|wind increasing|wind out|squall, moderate|bills of laden wind|hauling little)\\b"

wind_speed_patterns_stem_variants <- "(?i)\\b(strong(er)?|power(ful)?|heav(y|ier)?|intens(e|ity)?|light|gentl(e|y)?|soft|calm|clam|still|moder(ate|ating)?|mederate|mode|steady|gust(y|ing)?|bluster(ed|ing|y)?|blow(y|ing)?|brisk|variabl(e|y)?|changeabl(e|y)?|baffl(e|ed|ing)?|fresh|pleasant|fine|clear|perfect|smooth|eas(e|ed|ing)?|squall(s|y|ing)?|rain\\s+squall(s|y)?|storm(y|ier)?|tempest(uous|uousness)?|windstorm(s)?|big\\s+storm(s)?|terrific\\s+storm(s)?|tremendous|gale(s)?|perfect\\s+gale(s)?|high\\s+wind(s)?|hard|rough|grow(ing)?\\s+rough|fresh\\s+breeze|strong\\s+breeze|smart\\s+breeze|heavy\\s+breeze|very\\s+fresh\\s+breeze|fresh\\s+blow\\s+good\\s+breeze|whole\\s+sail\\s+breeze|peak\\s+breeze|fresh\\s+trade\\s+wind(s)?|light\\s+air|very\\s+light\\s+air(\\s+of\\s+wind)?|dead\\s+calm|perfectly\\s+calm|becalm(ed|ing)?|no\\s+wind|not\\s+wind\\s+enough|die(d|s)?\\s+out|clear(ed)?\\s+up|clear(ed)?\\s+of\\s+wind|wind\\s+out|haze(d)?\\s+up|wind\\s+increas(e|ing)|increas(ing|ed)?|breeze(d)?\\s+up|breez(ing|ed)?|haul(ed|ing)?(\\s+down|\\s+little|\\s+out)?|blow(ing)?\\s+hard|blow(ing)?\\s+a\\s+gale|blew\\s+very\\s+heavy|fearful(ly)?|blowed\\s+fearful(ly)?|high|scant|hair|fesh|sair|bills\\s+of\\s+laden\\s+wind|fresh\\s+breeze,\\s+rough|heavy\\s+blow,\\s+rough|squall,\\s+moderate|very\\s+moderate,\\s+calm|baffling\\s+wind|baffling\\s+and\\s+moderate)\\b"

```

## Read-in updated wind-speed categorizations 

```{r beaufort patterns}


wind_speed_patterns <- read_excel("data/Beaufort Wind Cats.xlsx")


```

## Code transferred


## Create a data frame for wind

```{r wind speed knots and conversions}

journal_wind_df <- journal_wind_unique

## Frances check: what might we be leaving out?

# takes the Beaufort categorizations we have and makes a column for wind speed in knots, according to standard conversion for each beaufort level 

journal_wind_df <- journal_wind_unique |>
  mutate(
    wind_speed_knots = case_when(
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$twelve)), collapse = "|")) ~ 64,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$eleven)), collapse = "|")) ~ 60,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$ten)), collapse = "|")) ~ 52,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$nine)), collapse = "|")) ~ 44,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$eight)), collapse = "|")) ~ 37,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$seven)), collapse = "|")) ~ 30,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$six)), collapse = "|")) ~ 24,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$five)), collapse = "|")) ~ 19,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$four)), collapse = "|")) ~ 13,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$three)), collapse = "|")) ~ 9,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$two)), collapse = "|")) ~ 5,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$one)), collapse = "|")) ~ 2,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$zero)), collapse = "|")) ~ 1,
      TRUE ~ NA_real_
    )
  )

journal_wind_df <- journal_wind_df |>
  
  # Convert meteorological wind direction to mathematical wind direction and make a new column 
  
  mutate(
  wind_dir_math = 270 - wind_meteor) |>
  
  mutate(wind_dir_math_real = ifelse(wind_dir_math <= 0, wind_dir_math + 360, wind_dir_math)
    ) |>
  
  # Convert wind direction (degrees) into radians 
  
  mutate(wind_dir_rad = wind_dir_math_real * (pi/180)) |>
   
  # Convert wind speed and direction into u and v vector components  
   
  mutate(u_vector = wind_speed_knots*(cos(wind_dir_rad))) |>

  mutate(v_vector = wind_speed_knots*(sin(wind_dir_rad))) |>

  # Calculate the magnitude of each vector 
  
  mutate(vector_magnitude = sqrt((u_vector^2)+(v_vector^2)))
  
```

## Add wind conversions and convert to radians 

```{r wind conversions to radians}

journal_wind_df <- journal_wind_df |>
 
   #convert meteor wind to math wind 
  
  mutate(wind_dir_math = 270 - wind_meteor) |>
  
  #if math wind is ≤ 0, add 360 
  
  mutate(wind_dir_math_real = ifelse(wind_dir_math <= 0, wind_dir_math + 360, wind_dir_math)) |>
  
  #convert to radians 
  
  mutate(wind_dir_rad = wind_dir_math_real * (pi / 180))

 
```

## Calculate u and v vectors 

```{r calculate u and v vectors}
journal_wind_df <- journal_wind_df |>
  mutate(u_vector = wind_speed_knots*(cos(wind_dir_rad))) 

(journal_wind_df <- journal_wind_df |>
  mutate(v_vector = wind_speed_knots*(sin(wind_dir_rad)))) 

```

## Calculate the magnitude of each vector 

```{r calculate vector magnitude}

 journal_wind_df <- journal_wind_df |>
  mutate(vector_magnitude = sqrt((u_vector^2)+(v_vector^2))) 

```

## Create a new column to represent months as numeric values, excluding NA values

```{r create month numeric column}

journal_wind_df <- journal_wind_df |>
  filter(!is.na(month)) |>
  mutate(
    month_numeric = match(month, month.name)  # Match month names to numeric values
  )

```

# Creating a beaufort column 

```{r add beaufort column}

# create a beaufort column 

journal_wind_df <- journal_wind_df |>
  mutate(
    beaufort = case_when(
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$twelve)), collapse = "|")) ~ 12,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$eleven)), collapse = "|")) ~ 11,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$ten)), collapse = "|")) ~ 10,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$nine)), collapse = "|")) ~ 9,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$eight)), collapse = "|")) ~ 8,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$seven)), collapse = "|")) ~ 7,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$six)), collapse = "|")) ~ 6,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$five)), collapse = "|")) ~ 5,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$four)), collapse = "|")) ~ 4,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$three)), collapse = "|")) ~ 3,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$two)), collapse = "|")) ~ 2,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$one)), collapse = "|")) ~ 1,
      str_detect(str_to_lower(wind_speed_description),
                 str_c(str_to_lower(na.omit(wind_speed_patterns$zero)), collapse = "|")) ~ 0,
      TRUE ~ NA_real_
    )
  )

```


# Make a dataframe to view beaufort frequencies: How often did x beaufort level come up? 

```{r look at beaufort frequencies and levels}

# Group for cardinal representation 
beaufort_grouped <- journal_wind_df |>
  group_by(beaufort, wind_direction) |>
  summarize(frequency = n()) 

# Group for mathematical representation 
beaufort_grouped_math <- journal_wind_df |>
  group_by(beaufort, wind_dir_math_real) |>
  summarize(frequency = n()) 

```


# Plot a frequency 'heat map'

```{r beaufort heat map}

ggplot(beaufort_grouped_math, aes(x = wind_dir_math_real, y = beaufort, fill = frequency)) +
  geom_tile()

```

# Beaufort base r plots 

```{r}
freq_tbl <- with(journal_wind_df, 
                 table(beaufort, wind_direction))

freq_tbl

```

```{r beaufort frequency by wind direction}

freq_df <- as.data.frame(freq_tbl)  # columns: beaufort, wind_direction, Freq

ggplot(freq_df, aes(x = wind_direction, y = beaufort, fill = Freq)) +
  geom_tile(color = "white") +  # creates a grid / heatmap
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(
    title = "Beaufort Frequency by Wind Direction",
    x = "Wind Direction",
    y = "Beaufort",
    fill = "Count"
  ) +
  theme_minimal() +
  theme(
    text = element_text(size = 12),
    axis.text.x = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

```





# Questions about beaufort data: 

```{r seasonal beaufort}

journal_wind_df <- journal_wind_df |>
  mutate(
    season = case_when(
      month_numeric %in% c(12, 1, 2) ~ "Winter",
      month_numeric %in% c(3, 4, 5)  ~ "Spring",
      month_numeric %in% c(6, 7, 8)  ~ "Summer",
      month_numeric %in% c(9,10,11)  ~ "Fall",
    ),
    season = factor(season, levels = c("Winter", "Spring", "Summer", "Fall"))
  )

```



```{r}
beaufort_palette <- c(
  "#f7fbff", "#deebf7", "#c6dbef", "#9ecae1",
  "#6baed6", "#4292c6", "#2171b5", "#08519c",
  "#08306b", "#4d004b", "#810f7c"
)

library(scico)

scale_fill_scico(palette = "lajolla", direction = 1)
```

```{r}
journal_wind_df |>
  filter(!is.na(wind_direction), !is.na(beaufort)) |>
  mutate(beaufort = factor(beaufort, ordered = TRUE)) |>
  count(wind_direction, beaufort) |>
  ggplot(aes(x = wind_direction, y = n, fill = beaufort)) +
  geom_col(width = 0.75, color = "grey30") +
  scale_x_discrete(limits = c("N","NNE","NE","ENE","E","ESE","SE","SSE","S","SSW","SW","WSW","W","WNW","NW","NNW", "No wind")) +
  scale_fill_manual(values = beaufort_palette, drop = FALSE) +
  labs(
    title = "Wind Direction and Strength",
    subtitle = "Stacked frequency by Beaufort category",
    x = "Wind direction",
    y = "Number of observations",
    fill = "Beaufort scale"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_text(color = "black"),
    axis.title = element_text(face = "bold"),
    plot.title = element_text(face = "bold")
  )

# ggsave(filename = "test_histo.jpg", width = 30, height = 15)
```



```{r}
journal_wind_df |>
  filter(!is.na(wind_direction), !is.na(beaufort)) |>
  mutate(beaufort = factor(beaufort, ordered = TRUE)) |>
  count(wind_direction, beaufort) |>
  group_by(wind_direction) |>
  mutate(prop = n / sum(n)) |>
  ggplot(aes(x = wind_direction, y = prop, fill = beaufort)) +
  geom_col(width = 0.75, color = "grey30") +
  scale_x_discrete(limits = c("N","NNE","NE","ENE","E","ESE","SE","SSE","S","SSW","SW","WSW","W","WNW","NW","NNW", "No wind")) +
  scale_fill_manual(values = beaufort_palette, drop = FALSE) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "Wind Direction and Strength (Proportional)",
    subtitle = "100% stacked bar — shows composition regardless of sample size",
    x = "Wind direction",
    y = "Proportion of observations",
    fill = "Beaufort scale"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_text(color = "black"),
    axis.title = element_text(face = "bold"),
    plot.title = element_text(face = "bold")
  )

```
```{r}
journal_wind_df |>
  filter(!is.na(wind_direction), !is.na(beaufort)) |>
  mutate(beaufort = factor(beaufort, ordered = TRUE)) |>
  count(wind_direction, beaufort) |>
  ggplot(aes(x = wind_direction, y = n, fill = beaufort)) +
  geom_col(width = 0.75, color = "grey30") +
  facet_wrap(~ wind_direction, scales = "free_y", ncol = 4) +
  scale_fill_manual(values = beaufort_palette, drop = FALSE) +
  labs(
    title = "Wind Direction and Strength (Faceted)",
    subtitle = "Each direction has its own y-scale for better visibility",
    x = "Wind direction",
    y = "Number of observations",
    fill = "Beaufort scale"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold")
  )

```
```{r}

journal_wind_df |>
  filter(!is.na(wind_direction), !is.na(beaufort)) |>
  mutate(beaufort = factor(beaufort, ordered = TRUE)) |>
  count(wind_direction, beaufort) |>
  group_by(wind_direction) |>
  mutate(
    prop = n / sum(n),
    total = sum(n)
  ) |>
  ungroup() |>
  ggplot(aes(x = wind_direction, y = prop, fill = beaufort)) +
  geom_col(width = 0.75, color = "grey30") +
  geom_text(
    data = . |> distinct(wind_direction, total),
    aes(x = wind_direction, y = 1.02, label = total),
    inherit.aes = FALSE,   # IMPORTANT
    size = 3,
    fontface = "bold"
  ) +
  scale_x_discrete(
    limits = c("N","NNE","NE","E","ESE","SE","S","SSW","SW","W","WNW","NW","No wind")
  ) +
  scale_fill_manual(values = beaufort_palette, drop = FALSE) +
  scale_y_continuous(
    labels = scales::percent_format(),
    expand = expansion(mult = c(0, 0.12))
  ) +
  labs(
    title = "Wind Direction and Strength (Proportional)",
    subtitle = "100% stacked bar — shows composition regardless of sample size\nNumbers above bars = sample size",
    x = "Wind direction",
    y = "Proportion of observations",
    fill = "Beaufort scale"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_text(color = "black"),
    axis.title = element_text(face = "bold"),
    plot.title = element_text(face = "bold")
  )



```
```{r}
directions <- c(
  "N","W","S","E"
)

journal_wind_df |>
  filter(
    wind_direction %in% directions,
    !is.na(beaufort),
    !is.na(season)
  ) |>
  mutate(beaufort = factor(beaufort, ordered = TRUE)) |>
  count(season, wind_direction, beaufort) |>
  group_by(season, wind_direction) |>
  mutate(
    prop  = n / sum(n),
    total = sum(n)
  ) |>
  filter(total > 30) |> 
  ungroup() |>
  ggplot(aes(x = wind_direction, y = total, fill = beaufort)) +

  geom_col(width = 1, color = "grey30") +

  # # geom_text(
  #   data = . |> distinct(season, wind_direction, total),
  #   aes(x = wind_direction, y = 1.08, label = total),
  #   inherit.aes = FALSE,
  #   size = 2.8,
  #   fontface = "bold"
  # ) +

  scale_fill_manual(values = beaufort_palette, drop = FALSE) +
  scale_x_discrete(limits = directions) +
  scale_y_continuous(limits = c(0,800)
  ) +

  coord_polar(theta = "x", start = 7*pi/4, direction = -1) +

  facet_wrap(~ season, nrow = 2) +

  labs(
    title = "Seasonal Wind Direction and Strength",
    subtitle = "Each panel shows Beaufort composition by direction\nNumbers indicate seasonal sample size",
    fill = "Beaufort scale"
  ) +

  theme_minimal(base_size = 15) +
  theme(
    panel.grid.major = element_line(color = "grey85"),
    panel.grid.minor = element_blank(),
    axis.title = element_blank(),
    axis.text.x = element_text(size = 14, face = "bold"),
    axis.text.y = element_text(size = 14),
    plot.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold"),
    legend.position = "right"
  )

ggsave(filename = "beaufort_rose_season.jpg", width = 10, height = 12)

```
```{r}
journal_wind_df |>
  filter(
    !is.na(wind_direction),
    !is.na(beaufort),
    !is.na(season),
    wind_direction !="No wind",
  ) |>
  mutate(beaufort = factor(beaufort, ordered = TRUE)) |>
  count(season, wind_direction, beaufort) |>
  group_by(season, wind_direction) |>
  mutate(
    prop  = n / sum(n),
    total = sum(n)
  ) |>
  ungroup() |>
  ggplot(aes(x = wind_direction, y = prop, fill = beaufort)) +

  geom_col(width = 0.75, color = "grey30") +

  geom_text(
    data = . |> distinct(season, wind_direction, total),
    aes(x = wind_direction, y = 1.05, label = total),
    inherit.aes = FALSE,
    size = 3.1,
    fontface = "bold"
  ) +

  scale_x_discrete(
    limits = c("N","NNE","NE","E","ESE","SE","S","SSW","SW","W","WNW","NW","No wind")
  ) +
  scale_fill_manual(values = beaufort_palette, drop = FALSE) +
  scale_y_continuous(
    labels = scales::percent_format(),
    expand = expansion(mult = c(0, 0.15))
  ) +

  facet_wrap(~ season, ncol = 1) +

  labs(
    title = "Seasonal Wind Direction and Strength",
    subtitle = "100% stacked bars by season\nNumbers above bars = seasonal sample size",
    x = "Wind direction",
    y = "Proportion of observations",
    fill = "Beaufort scale"
  ) +

  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_text(color = "black", face = "bold"),
    axis.title = element_text(face = "bold"),
    plot.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold", size = 19, vjust = -.2)
  )

ggsave(filename = "beaufort_histo_season.jpg", width = 10, height = 12)
```


```{r}


directions <- c(
  "N","NNW","NW","WNW","W","WSW","SW","SSW","S","SSE","SE","ESE","E","ENE","NE","NNE"
)

journal_wind_df |>
  filter(
    wind_direction %in% directions,   # replaces != "No wind"
    !is.na(beaufort)
  ) |>
  mutate(beaufort = factor(beaufort, ordered = TRUE)) |>
  count(wind_direction, beaufort) |>
  group_by(wind_direction) |>
  mutate(
    prop  = n / sum(n),
    total = sum(n)
  ) |>
  ungroup() |>
  ggplot(aes(x = wind_direction, y = prop, fill = beaufort)) +

  # stacked proportional bars → wind rose wedges
  geom_col(
    width = 1,
    color = "grey30",
    position = "stack"
  ) +

  # total sample size annotation per direction
  geom_text(
    data = . |> distinct(wind_direction, total),
    aes(
      x = wind_direction,
      y = 1.08,
      label = total
    ),
    inherit.aes = FALSE,
    size = 3,
    fontface = "bold"
  ) +

  # KEEP the same Beaufort colors
  scale_fill_manual(values = beaufort_palette, drop = FALSE) +

  # IMPORTANT: lock compass ordering without factoring
  scale_x_discrete(limits = directions) +

  # proportions, with headroom for labels
  scale_y_continuous(
    labels = scales::percent_format(),
    limits = c(0, 1.1),
    expand = c(0, 0)
  ) +

  # wind rose orientation
  coord_polar(theta = "x", start = 29*pi/15, direction = -1) +

  labs(
    title = "Wind Direction and Strength (Wind Rose)",
    subtitle = "Stacked by Beaufort category\nNumbers indicate sample size per direction",
    fill = "Beaufort scale"
  ) +

  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major = element_line(color = "grey85"),
    panel.grid.minor = element_blank(),
    axis.title = element_blank(),
    axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 9),
    plot.title = element_text(face = "bold"),
    legend.position = "right"
  )



```



```{r}
beaufort_palette <- c(
  "#f7fbff", "#deebf7", "#c6dbef", "#9ecae1",
  "#6baed6", "#4292c6", "#2171b5", "#08519c",
  "#08306b", "#4d004b", "#810f7c"
)

library(scico)

scale_fill_scico(palette = "lajolla", direction = 1)


```


```{r}
journal_wind_df |>
  mutate(
    wind_direction = factor(
      wind_direction,
      levels = c("N", "NE", "E", "SE", "S", "SW", "W", "NW")
    )
  ) |>
  count(wind_direction) |>
  ggplot(aes(x = wind_direction, y = n)) +
  geom_col(width = 0.75) +
  geom_text(aes(label = n), vjust = -0.3, size = 3) +
  labs(
    title = "Frequency of Wind by Direction",
    subtitle = "All observations",
    x = "Wind direction",
    y = "Number of observations"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_text(color = "black"),
    axis.title = element_text(face = "bold"),
    plot.title = element_text(face = "bold")
  )

```


# How strong is the wind when it comes from each direction?
```{r}

journal_wind_df |>
  count(wind_direction, beaufort) |>
  ggplot(aes(x = wind_direction, y = beaufort, fill = n)) +
  geom_tile() +
  labs(
    title = "Wind Strength by Direction",
    x = "Wind direction",
    y = "Beaufort category",
    fill = "Count"
  ) +
  theme_minimal()

```

# How does wind strength change over time?
```{r}
ggplot(journal_wind_df, aes(x = month_numeric, y = as.numeric(beaufort))) +
  geom_violin(trim = FALSE) +
  labs(
    title = "Monthly Distribution of Wind Strength",
    x = "Month",
    y = "Beaufort scale"
  ) +
  theme_minimal()

```
```{r}
library(ggplot2)
library(scales)

journal_wind_df |>
  filter(!is.na(beaufort), !is.na(month_numeric)) |>
  mutate(
    month = factor(
      month_numeric,
      levels = 1:12,
      labels = month.abb
    )
  ) |>
  ggplot(aes(x = month, y = as.numeric(beaufort))) +

  # distribution
  geom_violin(
    trim = FALSE,
    fill = "grey80",
    color = "grey30",
    linewidth = 0.4
  ) +

  # central tendency (very important for interpretation)
  stat_summary(
    fun = median,
    geom = "point",
    size = 1.8,
    color = "black"
  ) +

  # optional IQR cue (subtle but helpful)
  stat_summary(
    fun.data = median_hilow,
    fun.args = list(conf.int = 0.5),
    geom = "linerange",
    linewidth = 0.8,
    color = "black"
  ) +

  scale_y_continuous(
    breaks = pretty_breaks(),
    expand = expansion(mult = c(0.02, 0.05))
  ) +

  labs(
    title = "Monthly Distribution of Wind Strength",
    subtitle = "Violin width indicates density; points show medians",
    x = "Month",
    y = "Beaufort scale"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 10)
  )

```




```{r}

journal_wind_df <- journal_wind_df |>
  dplyr::filter(
    date_mdy >= as.POSIXct("1871-01-01 00:00:00", tz = "UTC") &
    date_mdy <= as.POSIXct("1886-12-31 23:59:59", tz = "UTC")
  )


ggplot(journal_wind_df, aes(x = date_mdy, y = as.numeric(beaufort))) +
  geom_point(alpha = 0.25, size = 0.6) +
  labs(
    title = "Wind Strength Over Time",
    x = "Date",
    y = "Beaufort scale"
  ) +
  theme_minimal()

```

```{r}
ggplot(journal_wind_df, aes(x = date_mdy, y = as.numeric(beaufort))) +
  geom_point(alpha = 0.2, size = 0.5) +
  geom_smooth(
    method = "loess",
    span = 0.2,
    se = FALSE
  ) +
  labs(
    title = "Smoothed Wind Strength Over Time",
    x = "Date",
    y = "Beaufort scale"
  ) +
  theme_minimal()

```

```{r}
journal_wind_df |>
  mutate(year_month = floor_date(date_mdy, "month")) |>
  filter(date_mdy >= as.Date("1876-01-01") &
         date_mdy <= as.Date("1876-12-31")) |>
  group_by(year_month) |>
  summarise(
    mean_beaufort = mean(as.numeric(beaufort), na.rm = TRUE),
    sd_beaufort   = sd(as.numeric(beaufort), na.rm = TRUE),
    n = n()
  ) |>
  ggplot(aes(x = year_month, y = mean_beaufort)) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = mean_beaufort - sd_beaufort,
      ymax = mean_beaufort + sd_beaufort
    ),
    alpha = 0.2
  ) +
  labs(
    title = "Monthly Mean Wind Strength with Variability",
    x = "Date",
    y = "Mean Beaufort scale"
  ) +
  theme_minimal()



```
```{r}
journal_wind_df |>
  mutate(
    date_mdy = as.Date(date_mdy),   # <-- convert POSIX to Date
    year_month = floor_date(date_mdy, "month"),
    month = month(year_month, label = TRUE),
    season = case_when(
      month %in% c("Dec", "Jan", "Feb") ~ "Winter",
      month %in% c("Mar", "Apr", "May") ~ "Spring",
      month %in% c("Jun", "Jul", "Aug") ~ "Summer",
      TRUE ~ "Fall"
    )
  ) |>
  group_by(year_month, season) |>
  summarise(
    mean_beaufort = mean(as.numeric(beaufort), na.rm = TRUE),
    sd_beaufort   = sd(as.numeric(beaufort), na.rm = TRUE),
    .groups = "drop"
  ) |>
  ggplot(aes(x = year_month, y = mean_beaufort)) +
  geom_line(color = "black") +
  geom_ribbon(
    aes(
      ymin = mean_beaufort - sd_beaufort,
      ymax = mean_beaufort + sd_beaufort
    ),
    alpha = 0.2
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +
  labs(
    title = "Monthly Mean Wind Strength with Seasonal Context",
    x = "Month",
    y = "Mean Beaufort scale"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_text(face = "bold"),
    plot.title = element_text(face = "bold")
  )

ggsave(filename = "beaufort_time_series.jpg", width = 20, height = 6)

```


```{r}
journal_wind_df |>
  filter(
    !is.na(wind_direction),
    tolower(wind_direction) != "no wind"
  ) |>
  mutate(
    date_mdy = as.Date(date_mdy),
    year_month = floor_date(date_mdy, "month"),
    month = month(year_month, label = TRUE),
    season = case_when(
      month %in% c("Dec", "Jan", "Feb") ~ "Winter",
      month %in% c("Mar", "Apr", "May") ~ "Spring",
      month %in% c("Jun", "Jul", "Aug") ~ "Summer",
      TRUE ~ "Fall"
    )
  ) |>
  group_by(year_month, season) |>
  summarise(
    n_reports = n(),
    .groups = "drop"
  ) |>
  ggplot(aes(x = year_month, y = n_reports)) +
  geom_line(color = "black") +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +
  labs(
    title = "Frequency of Wind Direction Reports Over Time",
    x = "Month/Year",
    y = "Number of direction reports"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_text(face = "bold"),
    plot.title = element_text(face = "bold")
  )

ggsave(filename = "direction_time_series.jpg", width = 20, height = 6)
```
```{r}
journal_wind_df |>
  filter(
    !is.na(wind_speed_description),
    wind_speed_description != ""
  ) |>
  mutate(
    date_mdy = as.Date(date_mdy),
    year_month = floor_date(date_mdy, "month"),
    month = month(year_month, label = TRUE),
    season = case_when(
      month %in% c("Dec", "Jan", "Feb") ~ "Winter",
      month %in% c("Mar", "Apr", "May") ~ "Spring",
      month %in% c("Jun", "Jul", "Aug") ~ "Summer",
      TRUE ~ "Fall"
    )
  ) |>
  group_by(year_month, season) |>
  summarise(
    n_reports = n(),
    .groups = "drop"
  ) |>
  ggplot(aes(x = year_month, y = n_reports)) +
  geom_line(color = "black") +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +
  labs(
    title = "Frequency of Wind Speed Descriptions Over Time",
    x = "Month/Year",
    y = "Number of wind speed reports"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_text(face = "bold"),
    plot.title = element_text(face = "bold")
  )
ggsave(filename = "speed_time_series.jpg", width = 20, height = 6)
```



# Are some directions seasonally stronger/weaker?



# Is variability as important as central tendency?

```{r}

ggplot(journal_wind_df, aes(x = wind_direction, y = as.numeric(beaufort))) +
  geom_boxplot() +
  labs(
    title = "Variability of Wind Strength by Direction",
    x = "Wind direction",
    y = "Beaufort scale"
  ) +
  theme_minimal()

```
```{r}
directions <- c(
  "N","NNW","NW","WNW","W","WSW","SW","SSW","S","SSE","SE","ESE","E","ENE","NE","NNE"
)

ggplot(journal_wind_df, aes(x = wind_direction, y = as.numeric(beaufort))) +
  geom_boxplot() +
  scale_x_discrete(limits = directions) +   # orders directions properly
  labs(
    title = "Variability of Wind Strength by Direction",
    x = "Wind direction",
    y = "Beaufort scale"
  ) +
  theme_minimal()

```


```{r}
directions <- c(
  "N","NNW","NW","WNW","W","WSW","SW","SSW","S","SSE","SE","ESE","E","ENE","NE","NNE"
)

ggplot(journal_wind_df, aes(x = wind_direction, y = as.numeric(beaufort))) +
  geom_boxplot() +
  scale_x_discrete(limits = directions) +   # orders directions properly
  facet_wrap(~ season) +
  labs(
    title = "Directional Wind Variability by Season",
    x = "Wind direction",
    y = "Beaufort scale"
  ) +
  theme_minimal()

ggsave(filename = "beaufortbox.jpg", width = 15, height = 6)

```
```{r}
directions <- c("N", "W", "S", "E")

wind_common <- journal_wind_df |>
  filter(
    wind_direction %in% directions,
    !is.na(beaufort),
    !is.na(season)
  ) |>
  group_by(season, wind_direction) |>
  mutate(total = n()) |>
  filter(total > 30) |>   # SAME threshold as rose
  ungroup()

```

```{r}
ggplot(
  wind_common,
  aes(
    x = factor(wind_direction, levels = directions),
    y = as.numeric(beaufort)
  )
) +
  geom_boxplot(
    width = 0.7,
    outlier.alpha = 0.4
  ) +
  facet_wrap(~ season, nrow = 2) +
  labs(
    title = "Directional Wind Variability by Season",
    subtitle = "Boxplots use the same directions and datapoints as the wind rose",
    x = "Wind direction",
    y = "Beaufort scale"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(face = "bold"),
    plot.title = element_text(face = "bold")
  )

ggsave(filename = "beaufortbox.jpg", width = 15, height = 6)
```



```{r}


ggplot(
  journal_wind_df |> 
    filter(!is.na(wind_direction), !is.na(beaufort)),
  aes(x = wind_dir_math_real, color = factor(beaufort))
) +
  geom_density() +
  scale_x_continuous(
    limits = c(0, 360),
    breaks = seq(0, 360, 45)
  ) +
  labs(
    x = "Wind direction (degrees)",
    y = "Density",
    color = "Beaufort"
  )


```





```{r}

# Count non-NA wind directions per Beaufort
valid_beaufort <- journal_wind_df |>
  filter(!is.na(wind_dir_math_real), !is.na(beaufort)) |>
  group_by(beaufort) |>
  summarise(n_dir = sum(!is.na(wind_dir_math_real))) |>
  filter(n_dir >= 2) |>
  pull(beaufort)

# Keep only those Beaufort levels
plot_df <- journal_wind_df |>
  filter(beaufort %in% valid_beaufort, !is.na(wind_dir_math_real))

ggplot(plot_df, aes(x = wind_dir_math_real)) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  facet_wrap(~ beaufort, ncol = 2) +
  scale_x_continuous(
    limits = c(0, 360),
    breaks = seq(0, 360, 45)
  ) +
  labs(
    x = "Wind direction (degrees)",
    y = "Density",
    title = "Directional density by Beaufort"
  )


```





```{r}


ggplot(
  journal_wind_df |>
    filter(!is.na(wind_direction), !is.na(beaufort), wind_direction != "No wind") |>
    mutate(wind_direction = factor(wind_direction, levels = c("N","NE","E","SE","S","SW","W","NW"))),
  aes(x = factor(beaufort), fill = factor(beaufort))
) +
  geom_bar() +
  facet_wrap(~ wind_direction, scales = "free_y") +  # one panel per direction
  labs(
    title = "Beaufort Distribution by Wind Direction",
    x = "Beaufort",
    y = "Count",
    fill = "Beaufort"
  ) +
  scale_fill_brewer(palette = "Spectral") +
  theme_minimal() +
  theme(
    text = element_text(size = 12),
    strip.text = element_text(face = "bold"),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )


```



# How many instances of directional data with no speed?

We could possibly go back in and see if we've missed anything on speed.

```{r}

# Rows 67 - 78 have NA wind speeds. Subtract Row 71: No Wind. 

sum(beaufort_grouped$frequency[67:78]) - beaufort_grouped$frequency[71]

```

# Create a dataframe pulling rows with mentioned wind direction but no speed

```{r}

missing_speed <- subset(journal_wind_df, !is.na(wind_direction) & is.na(wind_speed_knots))
view(missing_speed)

```

# Checking my math for distrubution table 
```{r}

sum(beaufort_grouped$frequency[1:65])

# Doesn't match, however not counting intercardinals yet (NNE, ENE, etc.)

```

# Pulling lines that have multiple wind descriptions recorded 

```{r}

wind_description_multiple <- journal_wind_df |>
  filter(!is.na(journal_wind_df$wind_speed_description),
         wind_speed_description != "",
         str_detect(wind_speed_description, "\\s"))
         

```

# Make a dataframe from main (journal_wind_df) showing rows with mentioned wind = yes only. Using to investigate "hauled" issue 

```{r}

yes_wind_mentioned <- journal_wind_df |>
  filter(wind_mentioned == c("Yes"))

# Show only entries with "hauled"

hauled_yes_wind_mentioned <- journal_wind_df |>
  filter(
    wind_mentioned == "Yes",
    str_detect(journal_entry, regex("hauled", ignore_case = TRUE))
  )

# Show me rows with direction and speed 

hauled_wind_yes_direction <- hauled_yes_wind_mentioned  |>
  filter(wind_direction != "No Wind", ignore_case = TRUE)

hauled_wind_yes_direction <- hauled_wind_yes_direction  |>
  filter(wind_direction != "No wind", ignore_case = TRUE)

# Write this dataframe to excel - will use generative AI to determine where he's talking about wood and where he's talking about wind 

write_xlsx(hauled_wind_yes_direction, "hauled_wind_yes_direction.xlsx")

# Find where he's explicitly saying "wind hauled"

wind_hauling <- journal_wind_df |>
  filter(grepl("wind hauled", journal_entry, ignore.case = TRUE))

# # View them
# wind_hauling$journal_entry




```

```{r}
hauled_wind_yes_direction$journal_entry[c(1, 2, 3, 4, 5,6,7,8,9,10)]

```

# Where are we getting beaufort and where are we not?
```{r}
beaufort_df <- journal_wind_df |>
  filter(!is.na(beaufort))

no_beaufort_df <- journal_wind_df |>
  filter(is.na(beaufort))

df_direction_words <- journal_wind_df |>
  filter(str_detect(journal_entry, "\\b(north|south|east|west)(?:ward)\\b"))


```


# Visualizing wind without speed

```{r wind without speed}


missing_speed <- subset(
  journal_wind_df,
  is.na(wind_speed_knots) &
    !(tolower(wind_direction) %in% "no wind")
)

missing_speed$wind_speed_fake <- 1


windRose(
  missing_speed,
  wd = "wind_meteor",
  ws = "wind_speed_fake",
  paddle = FALSE
)



```




```{r}

freq <- missing_speed |>
  count(month, wind_direction)

ggplot(freq, aes(wind_direction, month, fill = n)) +
  geom_tile() +
  coord_polar()

```

```{r}


month_levels <- c(
  "January", "February", "March", "April", "May", "June",
  "July", "August", "September", "October", "November", "December"
)

directions <- c(
  "N","W","S","E"
)

freq <- missing_speed |>
  mutate(month = factor(month, levels = month_levels),
         wind_direction = factor(wind_direction, levels = directions),
         season = season) |>
   filter(
    !is.na(wind_direction),
  ) |>
  count(month, wind_direction, season) |>
  complete(month, wind_direction, fill = list(n = 0)) |>
  mutate(n_fill = ifelse(n == 0, NA, n))

         
ggplot(freq, aes(wind_direction, month, fill = n_fill)) +
  geom_tile(color = "grey80") +
  scale_fill_gradient(
    low = "lightblue",
    high = "navy",
    na.value = "white",
    name = "Count"
  ) +
  scale_y_discrete(limits = rev(month_levels)) +
  labs(
    title = "Missing Wind Speed Observations by Season"
  ) +
  facet_wrap(~ season, nrow = 2) +
  
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major = element_line(color = "grey85"),
    panel.grid.minor = element_blank(),
    axis.title = element_blank(),
    axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 9),
    plot.title = element_text(face = "bold"),
    legend.position = "right"
  ) +
  coord_polar(theta = "x", start = 7*pi/4, direction = -1)

# ,"WNW","ESE","SSW","SSE"

ggsave(filename = "missing_wind_season.jpg", width = 10, height = 12)
```
```{r}
freq2 <- freq |>
  mutate(
    month_num = as.integer(month)
  )

```

```{r}

ggplot(freq, aes(
  x = wind_direction,
  y = month,
  fill = n_fill
)) +

  geom_col(
    width = 1,
    color = "grey30"
  ) +

  scale_fill_gradient(
    low = "lightblue",
    high = "navy",
    na.value = "white",
    name = "Count"
  ) +

  scale_y_discrete(limits = rev(month_levels)) +

  scale_y_continuous(
    breaks = seq_along(month_levels),
    labels = month_levels
  ) +

  coord_polar(theta = "x", start = 29*pi/15, direction = -1) +
  geom_text(
  data = dir_totals,
  aes(
    x = wind_direction,
    y = 85,
    label = total
  ),
  inherit.aes = FALSE,
  size = 3,
  fontface = "bold"
)+


  labs(
    title = "Missing Wind Speed Observations by Direction and Month",
    subtitle = "Radial bands represent months; color indicates number of missing records"
  ) +

  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major = element_line(color = "grey85"),
    panel.grid.minor = element_blank(),
    axis.title = element_blank(),
    axis.text.x = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 9),
    plot.title = element_text(face = "bold"),
    legend.position = "right"
  )

```
```{r}
dir_totals <- freq |>
  group_by(wind_direction) |>
  summarise(total = sum(n), .groups = "drop")

```


# Interesting visual I made accidentally 

```{r}
month_levels <- c(
  "January", "February", "March", "April", "May", "June",
  "July", "August", "September", "October", "November", "December"
)

directions <- c(
  "N","NNE","NE","ENE","E","ESE","SE","SSE","S","SSW","SW","WSW","W","WNW","NW","NNW"
)

freq <- missing_speed |>
  mutate(month = factor(month, levels = month_levels),
         wind_direction = factor(wind_direction, levels = directions)) |>
  count(month, wind_direction) |>
  complete(month, wind_direction, fill = list(n = 0)) |>
  mutate(n_fill = ifelse(n == 0, NA, n))

         
ggplot(freq, aes(wind_direction, month, fill = n_fill)) +
  geom_tile(color = "grey80") +
  scale_fill_gradient(
    low = "lightblue",
    high = "navy",
    na.value = "white",
    name = "Count"
  ) +
  scale_y_discrete(limits = rev(month_levels)) +
  labs(
    title = "Missing Speed Data Counts by Month",
    x = "Wind Direction",
    y = "Month"
  )
  coord_polar(theta = "x", start = 29*pi/15, direction = -1)
```

# Trying that same wind rose for the data we do have...need to make a new dataframe that just contains speed (could use UpSet for this)

```{r}

present_speed <- subset(journal_wind_df, !is.na(wind_direction) & !is.na(wind_speed_knots))


month_levels <- c(
  "January", "February", "March", "April", "May", "June",
  "July", "August", "September", "October", "November", "December"
)

directions <- c(
  "N","NNE","NE","ENE","E","ESE","SE","SSE","S","SSW","SW","WSW","W","WNW","NW","NNW"
)

freq_wind_present <- present_speed |>
  mutate(month = factor(month, levels = month_levels),
         wind_direction = factor(wind_direction, levels = directions)) |>
  count(month, wind_direction) |>
  complete(month, wind_direction, fill = list(n = 0)) |>
  mutate(n_fill = ifelse(n == 0, NA, n))  |>
  filter(!is.na(wind_direction))

         
ggplot(freq_wind_present, aes(wind_direction, month, fill = n_fill)) +
  geom_tile(color = "grey80") +
  scale_fill_gradient(
    low = "lightblue",
    high = "navy",
    na.value = "white",
    name = "Count"
  ) +
  scale_y_discrete(limits = rev(month_levels)) +
  labs(
    title = "Present Speed Data Counts by Month",
    x = "Wind Direction",
    y = "Month"
  ) +
  coord_polar(theta = "x", start = 29*pi/15, direction = -1)

```


```{r}
wind_rose <- missing_speed |>
  count(wind_direction)

```

```{r}
directions <- c("N","NE","E","ESE","SSE","S","SSW","SW","W","WNW","NW")

wind_rose$wind_direction <- factor(
  wind_rose$wind_direction,
  levels = directions
)

```

```{r}
ggplot(wind_rose, aes(wind_direction, n, fill = n)) +
  geom_col(width = 1) +
  coord_polar(start = pi/2, direction = -1) +
  theme_minimal()

```
```{r}
ggplot(freq, aes(wind_direction, n, fill = n)) +
  geom_col(width = 1) +
  coord_polar(start = pi/2, direction = -1) +
  facet_wrap(~ month)

```



```{r}
ggplot(missing_speed, aes(month, wind_direction, color = wind_direction)) +
  geom_point(size = 1)

```
# Working on UpSetR Plots. I was working with ComplexUpSet but really want UpSetR to work. 

# Create set columns - seasonal UpSet for winter 

```{r}


upset_df <- journal_wind_df |>
  mutate(
    WindMentioned = wind_mentioned == "Yes",
    HasWindDescription = !is.na(wind_speed_description) &
                         wind_speed_description !="",
    HasDirection  = !is.na(wind_dir_math_real),
    StrongWind    = beaufort %in% c(8,9,10),
    ModerateWind  = beaufort %in% c(5, 6, 7), 
    CalmWind      = beaufort %in% c(2,3,4),
    WindOut     = beaufort %in% c(0,1),
    Winter        = month_numeric %in% c(12, 1, 2),
    Spring        = month_numeric %in% c(3, 4, 5),
    Summer       = month_numeric %in% c(6, 7, 8),
    Fall          = month_numeric %in% c(9, 10, 11),
    year_numeric  = as.numeric(format(date_mdy, "%Y"))
  ) |>
  # keep numeric columns for attribute plots
  select(
    HasWindDescription, HasDirection, WindMentioned, StrongWind, ModerateWind, CalmWind, WindOut, Winter, Spring, Summer, Fall, beaufort, wind_meteor, month_numeric, wind_speed_knots, year_numeric, vector_magnitude
  )

# there are some outliers in the year data (translation mistakes) so filter these out or they will skew the axis 

upset_df <- upset_df |>
  filter(year_numeric >= 1500, year_numeric <= 1900)

# make month column a factor so axis ticks are whole numbers 

upset_df <- upset_df |>
  mutate(
    month_numeric = factor(
      month_numeric,
      levels = 1:12,
      labels = 1:12
    )
  )


```

# Make sure columns that populate as factors are numeric 

```{r}

# right now only the month_numeric column does this but if you change the columns to include in the upset_df you may need to write additional lines for those if they populate as integers or if you make them factors like I did above  

upset_df$month_numeric <- as.numeric(upset_df$month_numeric)

```

# Make sure the dataframe is being recognized as a dataframe and not a tibble

```{r}
upset_df <- as.data.frame(upset_df)

```

# Remove NAs 

```{r}
upset_df_clean <- upset_df |>
  filter(!is.na(beaufort),
         !is.na(wind_meteor),
         !is.na(month_numeric),
         !is.na(wind_speed_knots),
         !is.na(vector_magnitude))

```

# Recognize as data frame 

```{r}

# might be redundant could remove later 

upset_df_clean <- as.data.frame(upset_df_clean)
```

# Make logical sets numeric (i.e. 1 or 0 in place of TRUE/FALSE)

```{r}
sets <- c("HasWindDescription", "HasDirection", "WindMentioned", "StrongWind", "ModerateWind", "CalmWind", "WindOut",  "Winter", "Spring", "Summer", "Fall")

upset_df_clean2 <- upset_df_clean |>
  mutate(across(all_of(sets), ~ as.integer(.)))

```

# Make sure sets are read first 

```{r}
upset_df_clean2 <- upset_df_clean2 |>
  select(all_of(sets), everything())

```

# Recognize as data frame 

```{r}

# might be redundant could remove later 

upset_df_clean2 <- as.data.frame(upset_df_clean2)

```


# Plot 

```{r}

jpeg("UpSetWinter.jpg", width = 11, height = 10, units = "in", res = 300)

upset(
  upset_df_clean2,
  sets = sets,
  order.by = "freq",
  attribute.plots = list(
    gridrows = 60,
    plots = list(
      list(plot = scatter_plot, x = "year_numeric", y = "beaufort", col.by = "sets"),
      list(plot = scatter_plot, x = "month_numeric", y = "beaufort", col.by = "sets"),
      list(plot = scatter_plot, x = "beaufort", y = "vector_magnitude", col.by = "sets"),
      list(plot = scatter_plot, x = "wind_meteor", y = "vector_magnitude", col.by = "sets")
    ),
    ncols = 2
   )
  )
 



dev.off()



```





# ComplexUpset code if needed later. Can remove when UpSetR fig is finalized. 

```{r}

journal_upset_df <- journal_wind_df |>
  mutate(
    # ---- Sets (logical only) ----
    WindMentioned = wind_mentioned == "Yes",
    HasDirection  = !is.na(wind_dir_math_real),
    StrongWind    = !is.na(beaufort) & beaufort >= 3,
    Winter        = month_numeric %in% c(12, 1, 2),

    # ---- Numeric versions (NEW columns) ----
    wind_dir_deg  = suppressWarnings(as.numeric(as.character(wind_dir_math_real))),
    month_num     = suppressWarnings(as.numeric(as.character(month_numeric))),
    year_num      = as.numeric(lubridate::year(date_mdy))
  )

ggsave(filename = "upset.jpg", width = 30, height = 15)



```

# ComplexUpset code if needed later. Can remove when UpSetR fig is finalized. 

```{r}
upset(
  journal_upset_df,
  
  # The sets we want intersections for
  intersect = c("WindMentioned", "HasDirection", "StrongWind", "Winter"),

  # Base annotation: counts of intersection sizes
  base_annotations = list(
  "Intersection size" = intersection_size()
),

  
  # Attribute plots
  annotations = list(
    
    # Numeric scatter plots
    "Direction vs Beaufort" =
      ggplot(mapping = aes(x = wind_dir_deg, y = beaufort)) +
      geom_point(alpha = 0.4, size = 0.8, na.rm = TRUE) +
      labs(x = "Wind direction (degrees)", y = "Beaufort scale"),
    
    "Month vs Direction" =
      ggplot(mapping = aes(x = month_num, y = wind_dir_math_real)) +
      geom_point(alpha = 0.4, size = 0.8, na.rm = TRUE) +
      labs(x = "Month", y = "Wind direction (degrees)"),
    
    "Year vs Beaufort" =
      ggplot(mapping = aes(x = year_num, y = beaufort)) +
      geom_point(alpha = 0.4, size = 0.8, na.rm = TRUE) +
      labs(x = "Year", y = "Beaufort scale"),
    
    # Categorical plot for compass directions
    "Direction distribution" =
      ggplot(mapping = aes(x = factor(wind_direction))) +
      geom_bar(na.rm = TRUE) +
      labs(x = "Wind direction (compass)", y = "Count")
  )
)

ggsave(filename = "upset.jpg", width = 30, height = 15)

```

# ComplexUpset code if needed later. Can remove when UpSetR fig is finalized. 

```{r}

upset(
  journal_upset_df,
  
  # The sets we want intersections for
  intersect = c("WindMentioned", "HasDirection", "StrongWind", "Winter"),

  # Base annotation: counts of intersection sizes
  base_annotations = list(
  "Intersection size" = intersection_size()
),

  annotations = list(

  "Direction vs Beaufort" =
    ggplot(mapping = aes(x = wind_dir_math_real, y = beaufort)) +
      geom_point(alpha = 0.5, size = 1.5, na.rm = TRUE) +
      labs(x = "Wind direction (°)", y = "Beaufort") +
      theme_minimal(base_size = 14),

  "Month vs Direction" =
    ggplot(mapping = aes(x = month_numeric, y = wind_dir_math_real)) +
      geom_point(alpha = 0.5, size = 1.5, na.rm = TRUE) +
      labs(x = "Month", y = "Wind direction (°)") +
      theme_minimal(base_size = 14),

  "Year vs Beaufort" =
    ggplot(mapping = aes(x = year, y = beaufort)) +
      geom_point(alpha = 0.5, size = 1.5, na.rm = TRUE) +
      labs(x = "Year", y = "Beaufort") +
      theme_minimal(base_size = 14),

  "Direction distribution" =
    ggplot(mapping = aes(x = factor(wind_direction))) +
      geom_bar(na.rm = TRUE) +
      labs(x = "Wind direction (compass)", y = "Count") +
      theme_minimal(base_size = 14) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))  # rotate labels

  )
)
```



## Plot with animation by year 

```{r}
library(ggplot2)
library(gganimate)
library(gifski)
library(lubridate)

test <- journal_wind_df |>
  drop_na(longitude.y, u_vector, latitude.y, v_vector) |> #drop rows containing missing values
  mutate(longitude.y = as.numeric(longitude.y), #create longitude on x-axis 
         latitude.y = as.numeric(latitude.y)) |> #create latitude axis 
  mutate(year = lubridate::year(lubridate::as_date(date_mdy))) 
  
p <-  test |>
ggplot() +
  geom_segment(aes(x = longitude.y, 
                   y = latitude.y, 
                   xend = longitude.y + u_vector, yend = latitude.y + v_vector, 
                   color = vector_magnitude), 
               arrow = arrow(type = "closed", length = unit(0.1, "inches")), alpha = 0.3) +
  theme_minimal() +
  labs(x = "Longitude", y = "Latitude") +
  ggtitle('Wind Direction and Strength — {closest_state}') +
  scale_color_viridis_c() +
  transition_states(year) # Use transition_manual to control frames manually

animate(p, renderer = gifski_renderer(), duration = 50, nframes = 500)

anim <- animate(
  p,
  renderer = gifski_renderer(),
  duration = 40,
  nframes = 400,
  width = 800,
  height = 600,
  res = 120
)

anim_save("wind_vectors_animation1.gif", animation = anim)


  
```
```{r}
library(ggplot2)
library(gganimate)
library(gifski)
library(lubridate)
library(viridis)

test <- journal_wind_df |>
  drop_na(longitude.y, latitude.y, u_vector, v_vector) |>
  mutate(
    longitude.y = as.numeric(longitude.y),
    latitude.y  = as.numeric(latitude.y),
    year = year(as_date(date_mdy)),
    u_plot = u_vector * 0.2,   # scale for visual clarity
    v_plot = v_vector * 0.2
  )

p <- ggplot(test) +
  geom_segment(
    aes(
      x = longitude.y,
      y = latitude.y,
      xend = longitude.y + u_plot,
      yend = latitude.y + v_plot,
      color = vector_magnitude
    ),
    arrow = arrow(type = "closed", length = unit(0.06, "inches")),
    linewidth = 0.3,
    alpha = 0.4
  ) +
  coord_fixed() +
  scale_color_viridis_c(
    name = "Wind strength",
    option = "C"
  ) +
  labs(
    title = "Wind direction and strength — {closest_state}",
    subtitle = "Arrows show direction wind is coming from; color indicates magnitude",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold")
  ) +
  transition_states(
    year,
    transition_length = 2,
    state_length = 1
  ) +
  ease_aes("linear")

animate(
  p,
  renderer = gifski_renderer(),
  duration = 40,
  nframes = 400
)

anim <- animate(
  p,
  renderer = gifski_renderer(),
  duration = 40,
  nframes = 400,
  width = 800,
  height = 600,
  res = 120
)

anim_save("wind_vectors_animation2.gif", animation = anim)


```
```{r}
bounds <- test |>
  summarise(
    xmin = min(longitude.y, na.rm = TRUE),
    xmax = max(longitude.y, na.rm = TRUE),
    ymin = min(latitude.y, na.rm = TRUE),
    ymax = max(latitude.y, na.rm = TRUE)
  )

pad <- 0.2  # degrees; adjust to taste

```

```{r}
p <- test |>
  ggplot() +
  geom_segment(
    aes(
      x = longitude.y,
      y = latitude.y,
      xend = longitude.y + u_vector,
      yend = latitude.y + v_vector,
      color = vector_magnitude
    ),
    arrow = arrow(type = "closed", length = unit(0.08, "inches")),
    alpha = 0.3,
    linewidth = 0.4
  ) +
  coord_fixed(
    xlim = c(bounds$xmin - pad, bounds$xmax + pad),
    ylim = c(bounds$ymin - pad, bounds$ymax + pad),
    expand = FALSE
  ) +
  scale_color_viridis_c(option = "C") +
  theme_minimal(base_size = 13) +
  labs(x = "Longitude", y = "Latitude") +
  ggtitle("Year {closest_state}") +
  transition_states(year)

```


# Trying to add day and month now # Laurie and Frances are here 

```{r}
library(ggplot2)
library(gganimate)
library(gifski)
library(lubridate)
library(dplyr)

test <- journal_wind_df |>
  drop_na(longitude.y, u_vector, latitude.y, v_vector) |>
  mutate(longitude.y = as.numeric(longitude.y), 
         latitude.y = as.numeric(latitude.y)) |>
  mutate(month = month(lubridate::as_date(date_mdy), label = FALSE)) |>
  mutate(year = year(lubridate::as_date(date_mdy))) |>
  mutate(month_year = case_when(month > 9 ~ paste0(year, month),
                                month < 10 ~ paste0(year, 0, month)))
         
# # Shift positions incrementally to simulate wind movement 1/2
# test <- journal_wind_df |>
#   group_by(frame) |>
#   mutate(
#     long_shifted = long + (u_vector * (frame / num_frames)), # Incrementally shift longitude
#     lat_shifted = lat + (v_vector * (frame / num_frames))    # Incrementally shift latitude
#   ) |>
#   ungroup()
# # this code works when this chunk is not here 2/2

p <- test |>
  #filter(longitude.y < -20 & latitude.y < 20) |> ## If you wanted to make a bounding box for the GOM
  ggplot() +
  geom_segment(aes(x = longitude.y, 
                   y = latitude.y, 
                   xend = longitude.y + u_vector, 
                   yend = latitude.y + v_vector, 
                   color = vector_magnitude), 
               arrow = arrow(type = "closed", length = unit(0.1, "inches")), 
               alpha = 0.3) +
  theme_minimal() +
  labs(x = "Longitude", y = "Latitude") +
  ggtitle('Month and Year: {closest_state}') +
  scale_color_viridis_c() +
  enter_fade() +
  ease_aes('linear') +
  transition_states(month_year, state_length = 1) # Use month and year for animation

animate(p, renderer = gifski_renderer(), duration = 120, nframes = 100)

```

# Trying to simulate wind movement 

```{r}
num_frames <- 500

test <- journal_wind_df |>
  drop_na(longitude.y, u_vector, latitude.y, v_vector) |>
  mutate(
    longitude.y = as.numeric(longitude.y), 
    latitude.y = as.numeric(latitude.y)
  ) |>
  mutate(month_year = factor(paste(year(lubridate::as_date(date_mdy)), 
                                   month(lubridate::as_date(date_mdy), label = TRUE), 
                                   sep = " "))) |>
  crossing(frame = seq_len(num_frames)) |> # Add frame sequence
  mutate(
    long_shifted = longitude.y + (u_vector * (frame / num_frames)), # Incrementally shift longitude
    lat_shifted = latitude.y + (v_vector * (frame / num_frames))    # Incrementally shift latitude
  )

p <- test |>
  ggplot() +
  geom_segment(aes(
    x = long_shifted, 
    y = lat_shifted, 
    xend = long_shifted + u_vector, 
    yend = lat_shifted + v_vector, 
    color = vector_magnitude
  ), 
  arrow = arrow(type = "closed", length = unit(0.1, "inches")), 
  alpha = 0.3) +
  theme_minimal() +
  labs(x = "Longitude", y = "Latitude") +
  ggtitle('Month and Year: {closest_state}') +
  scale_color_viridis_c() +
  transition_states(frame, state_length = 1) # Use frame for animation

animate(p, renderer = gifski_renderer(), duration = 50, nframes = num_frames)

```


# Plot without animation and save as a .jpg 

```{r}
library(ggplot2)
library(gganimate)
library(gifski)
library(lubridate)

test <- journal_wind_df |>
  drop_na(longitude.y, u_vector, latitude.y, v_vector) |>
  mutate(longitude.y = as.numeric(longitude.y),
         latitude.y = as.numeric(latitude.y)) |>
  mutate(year = lubridate::year(lubridate::as_date(date_mdy)))
  
test |>
ggplot() +
  geom_segment(aes(x = longitude.y, 
                   y = latitude.y, 
                   xend = longitude.y + u_vector, yend = latitude.y + v_vector, 
                   color = vector_magnitude), 
               arrow = arrow(type = "closed", length = unit(0.1, "inches")), alpha = 0.3) +
  theme_minimal() +
  labs(x = "Longitude", y = "Latitude") +
  scale_color_viridis_c() +
facet_wrap(~year)


ggsave(filename = "wind_facets.jpg", width = 14, height = 7)
  
```


```{r}

test <- journal_wind_df |>
  drop_na(longitude.y, u_vector, latitude.y, v_vector) |>
  mutate(longitude.y = as.numeric(longitude.y),
         latitude.y = as.numeric(latitude.y)) |>
  mutate(year = lubridate::year(lubridate::as_date(date_mdy))) |>
  filter(year == 1886)
  
  
test |>
ggplot() +
  geom_segment(aes(x = longitude.y, 
                   y = latitude.y, 
                   xend = longitude.y + u_vector, yend = latitude.y + v_vector, 
                   color = vector_magnitude), 
               arrow = arrow(type = "closed", length = unit(0.1, "inches")), alpha = 0.3) +
  theme_minimal() +
  labs(x = "Longitude", y = "Latitude") +
  scale_color_viridis_c() +
facet_wrap(~year)


ggsave(filename = "wind_facets.jpg", width = 14, height = 7)



```

```{r}

test <- journal_wind_df |>
  drop_na(longitude.y, u_vector, latitude.y, v_vector) |>
  mutate(longitude.y = as.numeric(longitude.y),
         latitude.y = as.numeric(latitude.y)) |>
  mutate(year = lubridate::year(lubridate::as_date(date_mdy))) |>
  filter(year == 1873)
  
  
test |>
ggplot() +
  geom_segment(aes(x = longitude.y, 
                   y = latitude.y, 
                   xend = longitude.y + u_vector, yend = latitude.y + v_vector, 
                   color = vector_magnitude), 
               arrow = arrow(type = "closed", length = unit(0.1, "inches")), alpha = 0.3) +
  theme_minimal() +
  labs(x = "Longitude", y = "Latitude") +
  scale_color_viridis_c() +
facet_wrap(~year)


ggsave(filename = "wind_facet_1873.jpg", width = 14, height = 7)

```


```{r}

test <- journal_wind_df |>
  drop_na(longitude.y, u_vector, latitude.y, v_vector) |>
  mutate(longitude.y = as.numeric(longitude.y),
         latitude.y = as.numeric(latitude.y)) |>
  mutate(year = lubridate::year(lubridate::as_date(date_mdy))) |>
  filter(year == 1874)
  
  
test |>
ggplot() +
  geom_segment(aes(x = longitude.y, 
                   y = latitude.y, 
                   xend = longitude.y + u_vector, yend = latitude.y + v_vector, 
                   color = vector_magnitude), 
               arrow = arrow(type = "closed", length = unit(0.1, "inches")), alpha = 0.3) +
  theme_minimal() +
  labs(x = "Longitude", y = "Latitude") +
  scale_color_viridis_c() +
facet_wrap(~year)


ggsave(filename = "wind_facet_1874.jpg", width = 14, height = 7)

```


```{r}

test <- journal_wind_df |>
  drop_na(longitude.y, u_vector, latitude.y, v_vector) |>
  mutate(longitude.y = as.numeric(longitude.y),
         latitude.y = as.numeric(latitude.y)) |>
  mutate(year = lubridate::year(lubridate::as_date(date_mdy))) |>
  filter(year == 1876)
  
  
test |>
ggplot() +
  geom_segment(aes(x = longitude.y, 
                   y = latitude.y, 
                   xend = longitude.y + u_vector, yend = latitude.y + v_vector, 
                   color = vector_magnitude), 
               arrow = arrow(type = "closed", length = unit(0.1, "inches")), alpha = 0.3) +
  theme_minimal() +
  labs(x = "Longitude", y = "Latitude") +
  scale_color_viridis_c() +
facet_wrap(~year)


ggsave(filename = "wind_facets.jpg", width = 14, height = 7)

```






# Fran working on plotting this from https://theoceancode.netlify.app/post/wind_vectors/
```{r}

library(tidyverse)
library(scales)

ggplot(data = journal_wind_df, aes(x = date, y = wind_speed_knots)) +
  # Here we create the wind vectors as a series of segments with arrow tips
    geom_segment(aes(x = date, xend = date + u_vector, y = 0, yend = v_vector), 
                 arrow = arrow(length = unit(0.15, 'cm')), size = 0.5, alpha = 0.7) +
  # I think adding points at the base of the vectors makes the figure easier to read
    geom_point(aes(x = date, y = 0), alpha = 0.5, size = 1) +
  # Changing the dates to better match the range shown on the x axis
    scale_x_date(labels = date_format('%Y-%m-%d'), breaks = date_breaks('4 days')) +
  # Change the y axis labels to make sense
    scale_y_continuous(breaks = wind_speed_knots, labels = as.character(abs(y_axis)/date)) +
  # Change the x and y axis labels
    labs(x = NULL, y = 'Wind Speed (m/s)') +
    coord_equal(ylim = c(min(y_axis/wind_scale), max(y_axis/wind_scale))) +
    theme(axis.text.x = element_text(angle = 20, hjust = 0.4, vjust = 0.5, size = 8),
          axis.text.y = element_text(size = 8))

```


## Time series plottin'

```{r}
require(ggplot2)
require(scales)
require(gridExtra)
require(lubridate)

## Ok so this creates each axis...

dat <- data.frame(datetime = ISOdate(1885,1:12,1) +
                     seq(0:23)*60*60, wind_speed = runif(24, 25, 75))

# test <- journal_wind_df |>

## create wind speed data
dat$ws <- runif(nrow(dat), 0 , 15 )

## create wind direction data
dat$wd <- runif(nrow(dat), 0 , 360 )

# define an end point for geom_segment
dat$x.end <- dat$datetime + minutes(60)

ggplot(data = dat, aes(x = datetime, y = pollutant)) + ##ok so this is what i'm not sure of, how do i put wind speed on the y axis?? bc right now these values are essentially random 
  geom_line() +
  geom_segment(data = dat,
               linewidth = 1,
               aes(x = datetime,
                   xend = x.end,
                   y = pollutant,
                   yend = wd),
               arrow = arrow(length = unit(0.5, "cm"))) +
  theme()
```
## Plot time series 

```{r}
library(ggplot2)
library(lubridate)

# use data from my data frame

journal_wind_df |>
  mutate(date_mdy = lubridate::ymd(date_mdy)) |>
  mutate(date_year = lubridate::year(date_mdy)) |>
  filter(date_year == 1880) |>
  #filter(date_mdy >= as.Date("1880-01-01") & date_mdy <= as.Date("1880-12-31")) |>
ggplot(aes(x = date_mdy, y = wind_speed_knots)) + # x-axis = date, y-axis = wind speed (wind speed over time graph ! )
  geom_point(color = "blue", size = 0.5, alpha = 0.5) +  # creating a scatter plot
  geom_segment(aes(
    xend = date_mdy + u_vector * 0.01,   # adjust for horizontal (u_vector) movement (0.05 scaling factor should i change this??)
    yend = wind_speed_knots + v_vector * 0.05),  # adjust for vertical (v_vector) movement 
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red") +  # wind direction arrows in red
#  scale_x_datetime(date_labels = "%Y-%m", date_breaks = "2 months") +  # format X-axis and label with year and month 
  labs(title = "Wind Vectors Over Time", x = "Year-Month", y = "Wind Speed (knots)") +
  theme_minimal() 

```
# Making my time series into a slow reveal 

```{r}
library(ggplot2)
library(lubridate)
library(gganimate)

# use data from my data frame

journal_wind_df |>
  mutate(date_mdy = lubridate::ymd(date_mdy)) |>
  mutate(date_year = lubridate::year(date_mdy)) |>
  # filter(date_year < 1950) |>
  # filter(date_mdy >= as.Date("1873-10-08") & date_mdy <= as.Date("1874-04-13")) |> # using arbitrary dates to test a smaller subset of data
ggplot(aes(x = date_mdy, y = wind_speed_knots)) + # x-axis = date, y-axis = wind speed (wind speed over time graph ! )
  geom_point(color = "blue", size = 0.5, alpha = 0.5) +  # creating a scatter plot
  geom_segment(aes(
    xend = date_mdy + u_vector * 0.01,   # adjust for horizontal (u_vector) movement (0.05 scaling factor should i change this??)
    yend = wind_speed_knots + v_vector * 0.05),  # adjust for vertical (v_vector) movement 
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red") +  # wind direction arrows in red
#  scale_x_datetime(date_labels = "%Y-%m", date_breaks = "2 months") +  # format X-axis and label with year and month 
  labs(title = "Wind Vectors Over Time", x = "Year-Month", y = "Wind Speed (knots)") +
  theme_minimal() +
  transition_reveal(date_mdy)

 animate
```


## Create a plot showing u and v components in a scatter

```{r}
library(ggplot2)
library(lubridate)
library(dplyr)


journal_wind_df |>
  mutate(date_mdy = ymd(date_mdy)) |>
  mutate(date_year = year(date_mdy), date_month = month(date_mdy, label = TRUE)) |>  # Extract year and month
  filter(date_year == 1873) |>  
  ggplot(aes(x = date_mdy, y = 0)) +  
  geom_point(color = "blue", size = 0.5, alpha = 0.5) +  
  geom_segment(aes(
    xend = date_mdy + u_vector * 0.75,   
    yend = v_vector * 0.75),  
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red") +  
  # scale_x_date(date_labels = "%b", date_breaks = "1 month") +  # Format X-axis to show months
  labs(title = "Wind Vectors Over Time (1873)", x = "Month", y = "Wind Speed (knots)") +
  theme_minimal()

journal_wind_df |>
  ggplot(aes(x= u_vector, y = v_vector)) + geom_point()
# journal_wind_df |>
#   mutate(date_mdy = ymd(date_mdy)) |>
#   mutate(date_year = year(date_mdy), date_month = month(date_mdy, label = TRUE)) |>  # Extract year and month
#   filter(date_year == 1873) |>  
#   mutate (date_xend = date_mdy + u_vector * 0.5) |> 
#   mutate (date_yend = v_vector * 0.3) |> 
#   select(date_xend, date_yend, date_mdy)




```

See this tutorial for adding a label to the plot: https://stackoverflow.com/questions/41728575/how-to-add-specific-text-for-each-frame-of-an-animation-in-r


```{r}
journal_wind_df |>
  mutate(date_mdy = lubridate::ymd(date_mdy)) |>
  mutate(date_year = lubridate::year(date_mdy),
         date_month = lubridate::month(date_mdy)) |>
  drop_na(wind_speed_knots) |>
  filter(date_year < 1950) |>
  group_by(date_year, date_month, wind_speed_knots) |>
  summarize(num_wind_reported = n()) |>
  mutate(year_month = paste(date_year, date_month, "15", sep = "-"))# |>
 # ggplot(aes(x = lubridate::ymd(year_month), y = num_wind_reported, group = wind_speed_knots, color = wind_speed_knots)) +
 # geom_line() # +
  #facet_wrap(~wind_speed_knots)
```

# New term starting to work on a time series that makes more sense - experimenting witg plotting magnitude and direction?
```{r}

library(ggplot2)
library(lubridate)

# use data from my data frame

journal_wind_df |>
  mutate(date_mdy = lubridate::ymd(date_mdy)) |>
  mutate(date_year = lubridate::year(date_mdy)) |>
  filter(date_year == 1880) |>
ggplot(aes(x = date_mdy, y = wind_speed_knots)) + # x-axis = date, y-axis = wind speed (wind speed over time graph ! )
  geom_point(color = "blue", size = 0.5, alpha = 0.5) +  # creating a scatter plot
  geom_segment(aes(
    xend = date_mdy + vector_magnitude * 0.01,   # adjust for horizontal (u_vector) movement (0.05 scaling factor should i change this??)
    yend = wind_speed_knots + wind_dir_math_real * 0.05),  # adjust for vertical (v_vector) movement 
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red") +  # wind direction arrows in red
#  scale_x_datetime(date_labels = "%Y-%m", date_breaks = "2 months") +  # format X-axis and label with year and month 
  labs(title = "Wind Vectors Over Time", x = "Year-Month", y = "Wind Speed (knots)") +
  theme_minimal() 

```

## Next year 

# working with https://theoceancode.netlify.app/post/wind_vectors/

```{r}
library(tidyverse)
library(scales)


journal_wind_df |>
  mutate(date_mdy = lubridate::ymd(date_mdy)) |>
  mutate(date_year = lubridate::year(date_mdy)) |>
  filter(date_year == 1880) |>
  ggplot(aes(x = longitude.y, y = latitude.y)) +
  geom_segment(aes(
    xend = longitude.y + u_vector * 0.025,
    yend = latitude.y + v_vector * 0.025
  ),
  arrow = arrow(length = unit(0.01, "cm")),
  size = 0.25
  ) +
  coord_fixed() +
  facet_wrap(~ date_mdy) +  
  theme_minimal()

```


```{r}
wind_scale <- 1
y_axis <- seq(-5, 5, 5)

ggplot(data = journal_wind_df, aes(x = date, y = y_axis)) +
  # Here we create the wind vectors as a series of segments with arrow tips
    geom_segment(aes(x = date, xend = date + u*wind_scale, y = 0, yend = v*wind_scale), 
                 arrow = arrow(length = unit(0.15, 'cm')), size = 0.5, alpha = 0.7) +
  # I think adding points at the base of the vectors makes the figure easier to read
    geom_point(aes(x = date, y = 0), alpha = 0.5, size = 1) +
  # Changing the dates to better match the range shown on the x axis
    scale_x_date(labels = date_format('%Y-%m-%d'), breaks = date_breaks('4 days')) +
  # Change the y axis labels to make sense
    scale_y_continuous(breaks = y_axis, labels = as.character(abs(y_axis)/wind_scale)) +
  # Change the x and y axis labels
    labs(x = NULL, y = 'Wind Speed (m/s)') +
    coord_equal(ylim = c(min(y_axis/wind_scale), max(y_axis/wind_scale))) +
    theme(axis.text.x = element_text(angle = 20, hjust = 0.4, vjust = 0.5, size = 8),
          axis.text.y = element_text(size = 8))
```


```{r}
library(ggplot2)
library(lubridate)

# use data from my data frame

journal_wind_df |>
  mutate(date_mdy = lubridate::ymd(date_mdy)) |>
  mutate(date_year = lubridate::year(date_mdy)) |>
  filter(date_year == 1880) |>
  #filter(date_mdy >= as.Date("1880-01-01") & date_mdy <= as.Date("1880-12-31")) |>
ggplot(aes(x = date_mdy, y = wind_speed_knots)) + # x-axis = date, y-axis = wind speed (wind speed over time graph ! )
  geom_point(color = "blue", size = 0.5, alpha = 0.5) +  # creating a scatter plot
  geom_segment(aes(
    xend = date_mdy + u_vector * 0.01,   # adjust for horizontal (u_vector) movement (0.05 scaling factor should i change this??)
    yend = wind_speed_knots + v_vector * 0.05),  # adjust for vertical (v_vector) movement 
    arrow = arrow(length = unit(0.2, "cm")), 
    color = "red") +  # wind direction arrows in red
#  scale_x_datetime(date_labels = "%Y-%m", date_breaks = "2 months") +  # format X-axis and label with year and month 
  labs(title = "Wind Vectors Over Time", x = "Year-Month", y = "Wind Speed (knots)") +
  theme_minimal() 

```




